---
title: "STAT332 Statistical Analysis of Networks Sec 01 - Project 1 - Prof. Joe Reid"
author: "Andrew Sparkes"
date: "`r Sys.Date()`"
due: "16-02-2026"
output:
  html_document:
    self_contained: true
    df_print: paged
    code_folding: show
  pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(igraph) # Our standard graph object.
library(readr) # Reading in the data
library(gbRd)
library(ggraph) # Plotting
library(tidygraph) # Managing our graph object
library(RColorBrewer) # Getting Colors
library(scales) # scaling factors
library(grid) # for unit()
library(ggrepel) # for stronger label repel
library(visNetwork) # for network graphs
library(ggplot2) # for ploots
library(igraphdata)
library(dplyr)
library(patchwork)
```

```{r user_configurable_settings}
# user config

```

```{r helper_functions}
print_igraph_attr <- function(graph) {
  print("--- Graph Attributes ------")
  print(graph_attr(graph))

  print("--- Vertex Attributes ------")
  print(vertex_attr(graph))

  print("--- Edge Attributes ------")
  print(edge_attr(graph))
}

# copied from homework:
#Note that we are specifying directed reachability for directed graphs
reachability_cohesion <- function(g, mode = c("undirected", "weak", "strong")) {
  mode <- match.arg(mode)
  n <- vcount(g)
  if (n <= 1) return(1)

  if (mode == "undirected") {
    g2 <- as_undirected(g, mode = "collapse",
                        edge.attr.comb = list(weight = "sum"))
    D <- distances(g2)
    reachable <- sum(D != Inf) - n      # exclude self-pairs
    total <- n * (n - 1)

  } else if (mode == "weak") {
    # treat as undirected for reachability
    comps <- components(g, mode = "weak")$membership
    comp_sizes <- table(comps)
    reachable <- sum(comp_sizes * (comp_sizes - 1))
    total <- n * (n - 1)

  } else if (mode == "strong") {
    D <- distances(g, mode = "out")
    reachable <- sum(D != Inf) - n
    total <- n * (n - 1)
  }

  cohesion <- reachable / total
  return(cohesion)
}
```



For your midterm project, I would like you use use the data from the `igraphdata` package. You need to use the US airport network, **2010 December** dataset (which has **755** nodes and **23,473** edges) to write your report. This network has a number of node and edge attributes.

You may augment the node/edge attributes if it will help you with this project, but please remember to reference all data sources.

## Task:

Write a report using the US airport network data.  This report must do three things:

1.) Tell a story (the analysis needs to be tied together to make a specific point).
2.) Have well designed and attractive network graphs to support the story.
3.) Use general topology and other network measures to support the story you are trying to tell.

You **DO NOT** need to use the _full_ network. You may subset or aggregate the data as needed to focus your story (for example, you might look only at airports in the northeast or major hubs only, etc...).

# Story

```{r load_in_dataset}
data("USairports")
```

> This graph was created by an old(er) igraph version."
> ℹ Call `igraph::upgrade_graph()` on it to use with the current igraph version.
> For now we convert it on the fly...

Well, what is this?!?! Our airport/flight data is old right out of the gate? Considering it was based on data collected around December 2010, I'd suppose maybe it'd make sense for the dataset to potentially be created originally with an older versions of `igraph`, but let's investigate this a little to determine if this is a something negligible or a real problem. This warning says that the original graph was created by an old version, but it then proceeds to say that `igraph` will convert it on the fly... So, I guess no need to call `igraph::upgrade_graph(USairports)` then and we can carry on despite the weird info/warning popping up.... but.... let's embrace the paranoia and confirm:

```{r could_it_be_true}
graph_version() == graph_version(USairports)
```

Well, okay then... our versions apparently _do_ match up! But this still triggers that informational... That's an exploration for another day. Now, onward!

Now, let's do what most people never do, and see what we can learn from the manual (!) by using the built-in help function `?`:

```{r}
?USairports
```

Reading what the help documentation says in the _Help_ pane, we learn all kinds of context and descriptions about the data in our graph. A brief summary of what we learn and what it means about our graph object:
- the graph represents a network of airports and passenger connecting flights compiled based on December 2010 data
- that the edges (flights) are directed (ie. the flights have a directed orientation from A to B)
- that each edge is specific to a single carrier aircraft type
- that multiple carriers between the same two airports are denoted by multiple edges (we're looking at a multigraph object)
- that the graph itself has a `name` attribute: `r igraph::get_graph_attr(g, "name")`
- that each vertex (airport) has the following attributes:
  - name: the three-letter IATA airport code
  - City: city and state where the airport is located
  - Position: the geographical/physical coordinates of the airport, in WGS coordinates
- that each edge (flight) has the following attributes:
  - Carrier: the name of the airline
  - Departures: the number of departures for a given airline and aircraft type
  - Seats: the _total_ number of seats available on flights carried out by a given airline on a given aircraft type
  - Passengers: the _total_ number of passengers on the flight carried out by a given airline on a given aircraft type
  - Aircraft: the type of the aircraft used for the flight
  - Distance: the distance between the source and destination airports, in miles
- the file also has some additional information about data collection and sources, but that information is largely irrelevant for our purposes.

So, let's distill that back into something a little more readable: we know some data about airports (_what_ they're named, _where_ they are located geographically, and _what_ city/state they're in) and flights between them (the airport coming _from_, the airport going _to_, the _carrier_, the number of _departures_ for airline/aircraft combo, the number of _seats_ and the number of _passengers_ on the flight, the _aircraft_ type, and the _distance_ between the origin and destination airports).

That's a fairly good start at a general grasp of what kind of data points we have and what they describe in the real world, and we can even deduce some graph, vertex and edge attributes from the dataset's help file text, but that is more qualitative, but what about the _quantitative_, the actual data? Let's actually load our dataset into our R environment (and give it a shorthand alias of `g` to save some keystrokes):

```{r load_in_data}
g <- USairports # rename/alias to g
rm(USairports) # remove the original from environment for less opportunity for confusion

# let's confirm we got our igraph object created by coercing it to a string (and then extracting more information about it):
g
```

Now that we actually have the dataset loaded in and aliased for ease of use, we should consider a high level overview of _graph_-scale information, then we'll look at _vertices_ and their attributes, and finally we'll look at the _edges_ and their attributes. Once we have a rough idea of what data is actually in our dataset and how its related and represented, we'll be better able to get a good direction to explore a subset and analyze various measures and metrics to learn about the relationships therein. But let's not get ahead of ourselves, let's get back to that string summary of our graph `g`, see what's next and what we can learn about it: 

> IGRAPH bf6202d DN-- 755 23473 -- US airports

First off, this line tells us a few (well, quite a few) things in a short number of characters. Right off, we know:
- it's an `igraph` object (as we expected)
- its first $7$ characters of the hexidecimal (unique) identifier for the graph object (`bf6202d`)
- its four-letter `igraph` code string, indicating we have a **D**irected, **N**amed, unweighted, and is not bipartite
- it has $755$ vertices (airports)
- it has $23,473$ edges (carrier flights)
- the `name` of the graph is `US airports`

Well now, that is a fair number of edges! We'll certainly have to subset this graph down to a more manageable subgraph to analyze, but let's continue extracting more information about the dataset from its string summary and look at the second line: 

> + attr: name (g/c), name (v/c), City (v/c), Position (v/c), Carrier (e/c), Departures (e/n), Seats (e/n), Passengers (e/n), Aircraft (e/n), Distance (e/n)

From this very information-dense line, we learn that the graph `g` has the following attributes:
  - graph has attribute `name`, a `chr` type
  - vertices have attribute `name`, a `chr` type
  - vertices have attribute `City`, a `chr` type
  - vertices have attribute `Position`, a `chr` type
  - edges have attribute `Carrier`, a `chr` type
  - edges have attribute `Departures`, a `numeric` type
  - edges have attribute `Seats`, a `numeric` type
  - edges have attribute `Passengers`, a `numeric` type
  - edges have attribute `Aircraft`, a `numeric` type
  - edges have attribute `Distance`, a `numeric` type

Now, let's try looking at the unique values that we could perhaps use to filter our dataset down to something more reasonable... Let's see how many unique carriers are represented:


```{r}
length(unique(E(g)$Carrier))
```

$118$ carriers, eesh, okay. Maybe let's tone that down later. But first, let's look at the biggest airports instead, that may be a better starting point.

```{r}
length(unique(V(g)$name))
```

$755$ airport connections listed in this, eesh, well. Let's see if we can subset these reasonably. Maybe the top 10? By departures outbound?

```{r}
departure_strengths <- strength(g, weights = V(g)$Departures)
V(g)$departure_strengths <- departure_strengths

top_10 <- head(sort(departure_strengths, decreasing = TRUE), n = 10) # pull top 10 sorted by connecting flights
top_10
```




```{r}
# we have the top 10, now let's use those three digit airport codes to filter down the network
keep_node_codes <- V(g)[V(g)$name %in% list(names(top_10))[[1]]] # pruning vertex indices per top 10 list

g <- induced_subgraph(g, keep_node_codes) # the proper way to steal the nodes and all relevant edges
g
```

10 nodes and over a thousand edges, but many are duplicate, let's see about fixing that...


```{r}
E(g)[[1]]
```

```{r}
# g_simp <- simplify(g, edge.attr.comb = list(Departures = "sum", Carrier = "ignore", Passengers = "sum", Seats = "ignore", Aircraft = "ignore", Distance = "ignore")) # simplify will collapse all duplicate edges down to a single edge
# g_simp

# glimpse(edge_attr(g))

edge_df <- igraph::as_data_frame(g, what = "edges")

# Group by endpoints or edge groups and aggregate attributes
collapsed_df <- edge_df %>%
  group_by(from, to, Carrier, Aircraft, Seats = "sum", Passengers = "sum") %>%
  #mutate(Seats_total = sum(Seats), Passengers_total= sum(Passengers)) %>%

# Rebuild the simplified graph
g_collapsed <- graph_from_data_frame(collapsed_df, directed = FALSE)

```


```{r}
# print_igraph_attr(g_simp)
```



```{r}
# investigating biggest carriers?

```


```{r}
# filter further by Carrier
# carrier <- "Delta Air Lines Inc."

# this line triggers, and I don't know why: TODO look into why this is broken
# > Error in `FUN()`:
# > ! Unknown/unambigous attribute combination specification.
#sg <- simplify(sg, edge.attr.comb = list(weight = "sum", type = "concatenate")) # remove multiple edges

# instead, opting for a bit of a hack thanks to Stack Overflow
# sg <- igraph::as_data_frame(g) %>%
#   #filter(Carrier == carrier) %>%
#   group_by(Carrier) %>%
#   mutate(Departures_total = sum(Departures), Passengers_total = sum(Passengers)) %>%
#   unique() %>%
#   graph_from_data_frame()
# 
# sg

```

Now, let's calculate some things:
```{r}
reachability_cohesion(g)
diameter(g)
get_diameter(g)



cat("There are",
    length(bridges(g)),
    "bridges in this graph, and there are",
    length(articulation_points(g)),
    "articulation points in this graph.")

articulation_points(g)
farthest_vertices(g)

edge_density(g, loops = FALSE)

cohesion(g)
vertex_connectivity(g)
reachability_cohesion(g)

transitivity(g, type = "globalundirected", isolates = "NaN")
assortativity_degree(g)




```

```{r}
leiden_g <- cluster_leiden(as_undirected(g),
                           objective_function = "modularity",
                           resolution = 1,
                           n_iterations = 6,
                           vertex_weights = NULL)
plot(leiden_g, g)
```








```{r calculate_the_k_core}
k_core_g <- coreness(g, mode = "all")
```

`k_core_g` is not a graph object. It records the k-index when the nodes are removed from the graph.

```{r setup_and_plot_k_core_subgraph}


V(g)$core <- k_core_g # storing the already calculated k-core in the vertices' core attr

set.seed(1) # static seed for repeatability FOR NOW
L <- layout_with_fr(g) # Fruchterman-Reingold layout

ks <- sort(unique(V(g)$core)) # sorted vector of distinct k-core values
ks <- ks[ks > 0]  # drop 0-core isolates

# okay, breaking this down, looks like we are mapping over our distinct k-core values (in order, per sort call above), expecting a list return type and performing an anonymous function that is long enough to where it probably should be a _named_ function. Let's poke at it and comment on what it's doing...
plots <- lapply(ks, function(k){
  keep <- V(g)$core >= k # for any k passed in, keep holds the bool vector of vertex indices with >=k core
  sg <- induced_subgraph(g, vids = V(g)[keep]) # making a new subgraph consisting of just the vertices that have >= current k-core (and any connecting edges between them)

  # subset the full-graph layout to this subgraph's vertices (keeps positions stable)
  idx <- as.integer(V(g)[keep])        # vertex ids in original graph
  Lk  <- L[idx, , drop = FALSE] # subsetting the fr layout with the "keep" vertex ids/indices, the entire "row", and passing drop=FALSE ensures we get a df back out

  # passing along our subgraph to ggraph, manual layout with our Lk values from above loc
  ggraph(sg, layout = "manual", x = Lk[,1], y = Lk[,2]) +
    
    # now, adding relatively transparent mid gray edges connecting the vertices
    geom_edge_link(alpha = 0.25, colour = "grey60") +
    
    # adding the vertices, colored based on their k-core value, static size of 2.5
    geom_node_point(aes(color = core), size = 2.5) +
    
    # adding text labels, pulled from name attr, static size of 3, with a vertical adjustment
    geom_node_text(aes(label = name),
                 size = 3,
                 vjust = 1.5) +
    
    # setting the color scale, option D is viridis (the green/blue gradient one), excellent choice, the end param sets the end of the color hue range from [0,1] to [0, 0.95]
    # calling scale_color_viridis_c(), noting the _c ending, confused me for a while now, and it just dawned on me after reading the ggplot2 docs that that _c doesn't mean color option c, it stands for continuous, as opposed to d for discrete, or apparently b for binned color maps... Today I learned a thing...
    scale_color_viridis_c(option = "C", end = 0.95) +
    
    # add a plot title, indicating the current k value, as well as the number of vertices and edges in the subgraph
    ggtitle(paste0(k, "-core (core ≥ ", k, "), n=", vcount(sg), ", m=", ecount(sg))) +
    
    # blank theme with just default font size set
    theme_void(base_size = 11) +
    
    # nullify the legend, and set the plot title to be smaller font face and bolded
    theme(legend.position = "none",
          plot.title = element_text(size = 10, face = "bold"))
  
  # no explicit return, so that ggraph is the return obj
})

# from what I am guessing, this will distribute the plots among a 2-col grid layout, for as many plots exist in ... well, plots. I'm relating this to patchwork or kinda like faceting behavior... edit 27 minutes later, when I notice what package wrap_plots is from, and didn't notice when copying its library call to top of file.... Argh!
wrap_plots(plots, ncol = 2)

```

My interpretation from this core result is that the top 10 airport hubs in the US are very tightly engrained together, which is an unsurprising result.


```{r plot_the_graph_data}

p <- ggraph::ggraph(g_simp, layout = "fr") +
  geom_edge_fan(
    aes(size = Passengers),
    alpha = 0.8,
    color = "#ff0000"
  ) +
  geom_node_point() +
  geom_node_text(
    aes(
      label = name
      ),
    repel = TRUE,
    size = 3
  ) +
  
  


  labs(
    title = "Simple Plot"
  ) +
  theme_graph()
p

```


```{r}
ggraph(g, layout = "kk") + # same manual layout
    
  
    geom_edge_fan(aes(edge_color = Carrier)) +
  
    # this line differs. static vertex size and color... Hmmmm...
    geom_node_point(color = "#2C7FB8"
                 ) +
    
    # identical
    geom_node_text(aes(label = name),
                 vjust = 1.5) +
    
    # title the plot with the k-shell value, k-core value, and the vertex count
    ggtitle("") +
    
    # blank theme with base font size 11
    theme_void(base_size = 11)
```


First, let's get a rough idea of what we are working with. Our dataset is an imported `igraph` object, spanning some $755$ airports as vertices, and $23,473$ edges representing directed flights between them, with vertex data containing attributes for the airport's standard 3-character identifier, 


Possible avenues of interest.
Average occupancy percentage for flights from ALSDKJFALFDJS to LAKSDJFLAKSDJFSDLK, passengers/seat_count
maybe subsetting to flights with capacity over xyz number and see the patterns of their routes

Comparing average occupancy for flights that are longer distance vs shorter distance, or east/west cost, etc

Most popular inbound location, outbound locations, min/max distance flight (deg/str stuff)




subset our graph into a few subgraphs and see if we can notice any interesting patterns



```{r visnetwork}
# # Build nodes
# nodes <- igraph::as_data_frame(g, what = "vertices") |>
#   tibble::rownames_to_column("id") |> # this is an important function call, looking ahead to visNetwork's obsession with int ids
#   mutate(
#     id    = as.character(id),
#     label = media,
#     value = audience.size,
#     group = type.label
#   )
# 
# # Shape mapping
# node_types <- sort(unique(nodes$group))
# shape_vals <- c("dot", "square", "triangle")[seq_along(node_types)]
# shape_map <- setNames(shape_vals, node_types)
# nodes$shape <- shape_map[nodes$group]
# 
# # Node colors (background)
# node_colors <- setNames(scales::hue_pal()(length(node_types)), node_types)
# nodes$color.background <- node_colors[nodes$group]
# nodes$color.border <- "#444444"
# 
# nodes$title <- paste("Media Name: ", nodes$media, "<br>Media Type: ", nodes$type.label, "<br>Audience Size: ", nodes$audience.size)
# 
# # Build edges
# edges <- igraph::as_data_frame(g, what = "edges") |>
#   mutate(
#     from   = as.character(from),
#     to     = as.character(to),
#     width  = scales::rescale(weight, to = c(0.5, 1.5)),
#     arrows = "to",
#     smooth = TRUE
#   )
# 
# # Edge colors
# edge_types <- sort(unique(edges$type))
# edge_colors <- setNames(scales::hue_pal()(length(edge_types)), edge_types)
# edges$color <- edge_colors[edges$type]
# 
# # Dash mapping (linetype analogue)
# edges$dashes <- edges$type == edge_types[2] # example: second type dashed
# 
# edges$title <- paste("Reference Type: ", edges$type)
```

```{r}
# vis <- visNetwork(nodes, edges,
#   main = "Media Networks Link to and Reference Each Other",
#   submain = "A network plot of networks.",
#   footer = "Data from: https://kateto.net/network-visualization.",
#   width = "100%"
# ) |>
#   visNodes(scaling = list(min = 5, max = 30)) |>
#   visEdges(smooth = list(enabled = TRUE, type = "dynamic")) |>
#   visOptions(
#     highlightNearest = TRUE,
#     nodesIdSelection = TRUE,
#     collapse = TRUE
#   ) |>
#   visInteraction(
#     dragNodes = TRUE,
#     dragView  = TRUE,
#     zoomView  = TRUE
#   ) |>
#   visPhysics(stabilization = TRUE)
# 
# vis
```

# Pretty Graphs

```{r make_pretty_graphs}
# TODO pretty graphs
```

# Various Measures about subnetwork

```{r render_measures_table}

# structural_measures <- tribble(
#   # Headers---------------------------------------------------------------------
#   ~Measure, ~Value,
#   
#   # Diameter measures-----------------------------------------------------------
#   
#   "Diameter", diameter(g, weights = NA), # g is undirected, connected and has weights, but need to ignore weights for diameter to make sense
#   "Diameter Path", get_diameter(g, weights = NA), # need to ignore weights for this to make sense
#   "Farthest Vertices", farthest_vertices(g, weights = NA), # need to ignore weights for this to make sense
#   "Diameter with Reciprocal Weights", diameter(g, weights = 1/E(g)$weight),
#   "Diameter Path with Reciprocal Weights", get_diameter(g, weights = 1/E(g)$weight),
#   "Farthest Vertices with Reciprocal Weights", farthest_vertices(g, weights = 1 / E(g)$weight), # does this measure even make any sense?
#   
#   # Conductance measures--------------------------------------------------------
#   
#   # Adhesion
#   "Adhesion", adhesion(g),
#   
#   # Vertex Connectivity
#   "Vertex Connectivity", vertex_connectivity(g),
#   
#   # Fiedler's value (second smallest eigenvalue)
#   "Fiedler's Value (λ₂)", lambda_2
# )
# 
# structural_measures |>
#   gt() |>
#   tab_header(
#     title = "Graph Structural Measures",
#     subtitle = "Structural-related metrics about the Caviar dataset"
#   ) |>
#   # data_color(
#   #   columns = everything(),   # columns to tint
#   #   fn = function(cols) {
#   #     category <- global_measures_table$Category
#   #     category_colors[category]
#   #   }
#   # ) |> # TODO fix colors eventually
#   fmt_number(columns = Value, decimals = 6, use_seps = TRUE) |> # special formatting for numeric values
#   tab_options(table.font.size = 12)

```

