---
title: "STAT332 Statistical Analysis of Networks Sec 01 - Project 1 - Prof. Joe Reid"
author: "Andrew Sparkes"
date: "`r Sys.Date()`"
due: "16-02-2026"
output:
  html_document:
    self_contained: true
    df_print: paged
  pdf_document: default
---

# STAT 332: Project 1

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```


For your midterm project, I would like you use use the data from the igraphdata package. You need to use the US airport network, 2010 December dataset (which has 755 nodes and 23473 edges) to write your report.  This network has a number of node and edge attributes.

You may augment the node/edge attributes if it will help you with this project, but please remember to reference all data sources.

Task:

Write a report using the US airport network data.  This report must do three things:

1.)  Tell a story (the analysis needs to be tied together to make a specific point)
2.)  Have well designed and attractive network graphs to support the story.
3.)  Use general topology and other network measures to support the story you are trying to tell.

You DO NOT need to use the full network.  You may subset or aggregate the data as neededto focus your story (for example, you might look only at airports in the northeast or major hubs only, etc...).


```{r setup}
library(igraph) # Our standard graph object.
library(readr) # Reading in the data
library(ggraph) # Plotting
library(tidygraph) # Managing our graph object
library(RColorBrewer) # Getting Colors
library(scales) # scaling factors
library(grid) # for unit()
library(ggrepel) # for stronger label repel
library(visNetwork) # for network graphs
library(ggraph) # for graphs
library(ggplot2) # for ploots
```
