---
title: "STAT332 Statistical Analysis of Networks Sec 01 - Homework 5 - Prof. Joe Reid"
author: "Andrew Sparkes"
date: "`r Sys.Date()`"
due: "27-02-2026"
output:
  html_document:
    self_contained: true
    df_print: paged
    code_folding: show
  pdf_document: default
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(igraph) # Our standard graph object.
library(readr) # Reading in the data
library(gbRd)
library(gt)
library(ggraph) # Plotting
library(tidygraph) # Managing our graph object
library(RColorBrewer) # Getting Colors
library(scales) # scaling factors
library(grid) # for unit()
library(ggrepel) # for stronger label repel
library(visNetwork) # for network graphs
library(ggplot2) # for ploots
library(igraphdata)

source("../graph.R")
```

```{r user_configurable_settings}
# user config
football_gml_file_path <- "football/football.gml"
football_txt_file_path <- "football/football.txt"

```

I've given you a gml file for this homework. This one happens to be college football teams.

Now, I really know little to nothing about college football, but I'm pretty sure there are "communities" of teams that should be able to be identified. There are also teams that everyone wants to play in order to increase their rankings (cores?) etc...

Using a collection of the methods from class, find the subgroups within this network that might indicate clusters of teams. :)

# 0.) Load the dataset.

```{r}
g <- igraph::read_graph(football_gml_file_path, format = "gml")
```

```{r}
g
print_formatted_graph_summary_table(g)
```


# 1.) Plot things out.

```{r plooot}
set.seed(1)
p <- ggraph(g, layout = "lgl") +
  geom_edge_fan(
    aes(),
    alpha = 0.5
  ) +
  
  geom_node_point(
    aes(
      size = value
      ),
    alpha = 0.8
  ) +

  geom_node_text(
    aes(
      label = label,
      color = "#
      ff0000"
      ),
    repel = TRUE,
    size = 4
  ) +
  labs(
    title    = "A simple ugly plot",
    subtitle = "A network plot of football teams",
  ) +
  guides(position = "none")
p
```

```{r prep_for_plooot}
nodes <- igraph::as_data_frame(g, what = "vertices") |>
  mutate(
    title = paste0(label," (",value,")\n"),
    size = value,
    name = label
    )

edges <- igraph::as_data_frame(g, what = "edges") |>
  mutate(
    # from   = as.character(from),
    # to     = as.character(to),
    width  = 2,
    smooth = TRUE
  )

edges$color <- "#000000"
nodes$title <- paste0(
  nodes$label,
  " (",
  nodes$value,
  ")"
)

```

```{r use_visnetwork_instead}
set.seed(1)
vis <- visNetwork(nodes, edges,
                  main = "Football Plot",
                  submain = "",
                  footer = "",
                  width = "100%"
  ) |>
  visIgraphLayout(layout = "layout_nicely") |>
  visNodes(scaling = list(min = 5, max = 30)) |>
  visEdges(smooth = list(enabled = TRUE, type = "dynamic")) |>
  visOptions(
    highlightNearest = TRUE,
    nodesIdSelection = TRUE,
    collapse = TRUE
  ) |>
  visLegend() |>
  visInteraction(
    dragNodes = TRUE,
    dragView  = TRUE,
    zoomView  = TRUE
  ) |>
  visPhysics(stabilization = TRUE)

vis # we'll fuddle with this more later
```


# 2.) Use multiple methods and create matrices that represent how much agreement or disagreement the methods have.

```{r leiden_hosen}
leiden_g <- cluster_leiden(g,
                           objective_function = "modularity",
                           weights = NA,
                           resolution = 1,
                           n_iterations = 6,
                           vertex_weights = NULL)
nodes$membership_leiden = leiden_g$membership
```

```{r}
louvain_g  <- cluster_louvain(g, NA, resolution = 1)

nodes$membership_louvain = louvain_g$membership
```

```{r}
# let's not...
# plot(leiden_g, g)
# plot(louvain_g, g)
```

```{r use_visnetwork_instead_redux}
set.seed(1)
vis <- visNetwork(nodes, edges,
                  main = "Football Plot - Leiden Clustering",
                  submain = "",
                  footer = "",
                  width = "100%"
  ) |>
  visIgraphLayout(layout = "layout_nicely") |>
  visNodes(scaling = list(min = 5, max = 30)) |>
  visEdges(smooth = list(enabled = TRUE, type = "dynamic")) |>
  visOptions(
    highlightNearest = TRUE,
    nodesIdSelection = TRUE,
    collapse = TRUE
  ) |>
  visClusteringByGroup(nodes$membership_leiden) |>
  visLegend() |>
  visInteraction(
    dragNodes = TRUE,
    dragView  = TRUE,
    zoomView  = TRUE
  ) |>
  visPhysics(stabilization = TRUE)

vis # we'll fuddle with this more later
```

# 3.) Find the cores.

```{r calculate_the_k_cores_and_plot}
show_k_core_plots(g) # from graph.R
```

```{r calculate_the_k_core_shells_and_plot}
# show_k_core_shell_plots(g) # from graph.R # TODO figure out why this is shitting the bed not finding a name attribute somewhere in geom_node_text()
```

# 4.) Find the cliques.

```{r}
clique_num(g)
```

```{r}
cliq_set <- cliques(g, min = clique_num(g))
cliq_set
```

```{r}
largest_cliques(g)
```

```{r}
# plot(g, cliq_set[[1]], cliq_set[[2]]) # this line confuzzles the confuzzle out of me...
```

There appear to be two cliques in the data, both of size 9. I would have thought it'd line up with the size of the divisions, or perhaps geographically explained. Apparently there are 12 divisions in this league of college football. 

# 5.) Explain the results.

**Answer:**
```{r}
# TODO explanation
```
