---
title: "STAT332 Statistical Analysis of Networks Sec 01 - Homework 5 - Prof. Joe Reid"
author: "Andrew Sparkes"
date: "`r Sys.Date()`"
due: "27-02-2026"
output:
  html_document:
    self_contained: true
    df_print: paged
    code_folding: show
  pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(igraph) # Our standard graph object.
library(readr) # Reading in the data
library(gbRd)
library(ggraph) # Plotting
library(tidygraph) # Managing our graph object
library(RColorBrewer) # Getting Colors
library(scales) # scaling factors
library(grid) # for unit()
library(ggrepel) # for stronger label repel
library(visNetwork) # for network graphs
library(ggplot2) # for ploots
library(igraphdata)
```

```{r user_configurable_settings}
# user config
football_gml_file_path <- "football/football.gml"
football_txt_file_path <- "football/football.txt"

```

```{r helper_functions}
print_igraph_attr <- function(graph) {
  print("--- Graph Attributes ------")
  print(graph_attr(graph))

  print("--- Vertex Attributes ------")
  print(vertex_attr(graph))

  print("--- Edge Attributes ------")
  print(edge_attr(graph))
}
```


I've given you a gml file for this homework. This one happens to be
college football teams.

Now, I really know little to nothing about college football, but I'm pretty sure there are "communities" of teams that should be able to be identified. There are also teams that everyone wants to play in order to increase their rankings (cores?) etc...

Using a collection of the methods from class, find the subgroups within this network that might indicate clusters of teams. :)

# 0.) Load the dataset.

```{r}
g <- igraph::read_graph(football_gml_file_path, format = "gml")
```

```{r}
g
```


# 1.) Plot things out.

```{r}
set.seed(1)
  p <- ggraph::ggraph(g, layout = "fr") +
    geom_node_text(
      aes(label = label),
      repel = TRUE,
      size = 3
    ) +
    geom_edge_fan(
      aes(),
      alpha = 0.8
    ) +
  
    labs(
      title = "Simple Plot"
    ) +
    theme_graph()
  p
```


# 2.) Use multiple methods and create matrices that represent how much agreement or disagreement the methods have.

```{r}
# TODO what methods exactly?


V(g)$deg <- degree(g, mode="all", loops = FALSE, normalized=FALSE)

```


# 3.) Find the cores.

```{r calculate_the_k_cores_and_plot}
k_core_g <- coreness(g, mode = "all")

V(g)$core <- k_core_g # storing the already calculated k-core in the vertices' core attr

set.seed(1) # static seed for repeatability FOR NOW
L <- layout_with_fr(g) # Fruchterman-Reingold layout

ks <- sort(unique(V(g)$core)) # sorted vector of distinct k-core values
ks <- ks[ks > 0]  # drop 0-core isolates

# okay, breaking this down, looks like we are mapping over our distinct k-core values (in order, per sort call above), expecting a list return type and performing an anonymous function that is long enough to where it probably should be a _named_ function. Let's poke at it and comment on what it's doing...
plots <- lapply(ks, function(k){
  keep <- V(g)$core >= k # for any k passed in, keep holds the bool vector of vertex indices with >=k core
  sg <- induced_subgraph(g, vids = V(g)[keep]) # making a new subgraph consisting of just the vertices that have >= current k-core (and any connecting edges between them)

  # subset the full-graph layout to this subgraph's vertices (keeps positions stable)
  idx <- as.integer(V(g)[keep])        # vertex ids in original graph
  Lk  <- L[idx, , drop = FALSE] # subsetting the fr layout with the "keep" vertex ids/indices, the entire "row", and passing drop=FALSE ensures we get a df back out

  # passing along our subgraph to ggraph, manual layout with our Lk values from above loc
  ggraph(sg, layout = "manual", x = Lk[,1], y = Lk[,2]) +
    
    # now, adding relatively transparent mid gray edges connecting the vertices
    geom_edge_link(alpha = 0.25, colour = "grey60") +
    
    # adding the vertices, colored based on their k-core value, static size of 2.5
    geom_node_point(aes(color = core), size = 2.5) +
    
    # adding text labels, pulled from name attr, static size of 3, with a vertical adjustment
    geom_node_text(aes(label = label),
                 size = 3,
                 vjust = 1.5) +
    
    # setting the color scale, option D is viridis (the green/blue gradient one), excellent choice, the end param sets the end of the color hue range from [0,1] to [0, 0.95]
    # calling scale_color_viridis_c(), noting the _c ending, confused me for a while now, and it just dawned on me after reading the ggplot2 docs that that _c doesn't mean color option c, it stands for continuous, as opposed to d for discrete, or apparently b for binned color maps... Today I learned a thing...
    scale_color_viridis_c(option = "C", end = 0.95) +
    
    # add a plot title, indicating the current k value, as well as the number of vertices and edges in the subgraph
    ggtitle(paste0(k, "-core (core ≥ ", k, "), n=", vcount(sg), ", m=", ecount(sg))) +
    
    # blank theme with just default font size set
    theme_void(base_size = 11) +
    
    # nullify the legend, and set the plot title to be smaller font face and bolded
    theme(legend.position = "none",
          plot.title = element_text(size = 10, face = "bold"))
  
  # no explicit return, so that ggraph is the return obj
})

# this will distribute the plots among a 2-col grid layout, for as many plots exist in ... well, plots. I'm relating this to patchwork or kinda like faceting behavior... edit 27 minutes later, when I notice what package wrap_plots is from, and didn't notice when copying its library call to top of file.... Argh!
wrap_plots(plots, ncol = 2)

```


# 4.) Find the cliques.

```{r}
clique_num(g)
```
```{r}
cliques(g, min = clique_num(g))

```

```{r}
largest_cliques(g)
```

There appear to be two cliques in the data, both of size 9. I would have thought it'd line up with the size of the divisions, or perhaps geographically explained. Apparently there are 12 divisions in this league of college football. 

# 5.) Explain the results.

```{r}

```
