---
title: "Stat 407 Community Detection Homework"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
link-citations: true
natbib: true
#csl: ieee.csl
#bibliography: NetworkReferences.bib
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I've given you a gml file for this homework. This one happens to be
college football teams.\
Now, I really know little to nothing about college football, but I'm
pretty sure there are "communities" of teams that should be able to be
identified. There are also teams that everyone wants to play in order to
increase their rankings (cores?) etc...

Using a collection of the methods from class, find the subgroups within
this network that might indicate clusters of teams. :)

```{r load_dataset}
# TODO load dataset
```


1.) Plot things out.
```{r}
# TODO plot football graph
```


2.) Use multiple methods and create matrices that represent how much
agreement or disagreement the methods have.
```{r}
# TODO football matrices
```


3.) Find the cores.
```{r}
# TODO find football k-cores
```


4.) Find the cliques
```{r}
# TODO find football cliques
```


5.) Explain the results.

**Answer:**
laksdjflkasdjfkds

## References
