---
title: 'A1 - Why no edges?!?! Where did they all go?!!'
author: "Andrew Sparkes"
date: "`r Sys.Date()`"
output:
  html_document:
  self_contained: true
df_print: paged
pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(echo = TRUE)

library(igraph) # TODO look into quietly = TRUE, warn.conflicts = FALSE, verbose = FALSE params in library()
library(networkdata)
library(ggplot2)
library(tidygraph)
library(ggraph)
library(visNetwork)
library(tibble)
library(purrr)
```


```{r user_configurable_settings}
gml_filename <- "Homework_3_filtered_data.gml"
```

Now, we have

```{r setup_environment_like_it_was_to_reproduce}
# reproduce the original g and nodes/edges from the original graph above, just in case we modified it inadvertently above
g <- igraph::read_graph(gml_filename, format = "gml")
data <- visNetwork::toVisNetworkData(g, idToLabel = FALSE)

nodes <- data$nodes
edges <- data$edges
```

```{r edge_check}
glimpse(edges)
```

```{r nodes_check}
glimpse(nodes)
```

Reading through `?visNetwork` and `?toVisNetworkData` learned me some new information: the parameters `nodes` and `edges` passed into `visNetwork()` require id (an int) in the `to` and `from` columns, and the edges that their dumb helper function `toVisNetworkData()` failed to do, despite throwing the flag to _not_ convert id to label... Let's confirm that suspicion, correct that and continue...

```{r confirm_breaking_the_thing}
d_with_flag <- visNetwork::toVisNetworkData(g, idToLabel = TRUE)
d_without_flag <- visNetwork::toVisNetworkData(g, idToLabel = FALSE)
```

```{r}
glimpse(d_with_flag$nodes)
glimpse(d_without_flag$nodes)
```

```{r}
glimpse(d_with_flag$edges)
glimpse(d_without_flag$edges)
```

Comparing the `data.frame`s of nodes and edges from with and without the `idToLabel` flag set. So the `idToLabel` flag parameter in `toVisNetworkData()` doesn't do what I thought it does. If it's `TRUE`, it will add another column to the nodes `data.frame` called `label` containing the id... It doesn't seem to _convert_ anything past that, ie. it doesn't rename the from/to attributes of the edges to match the id.

So in conclusion, the `idToLabel` parameter in `toVisNetworkData()` (if `TRUE`) simply copies the `id` (dbl) to a column named `label` (also dbl) as shown below in a toy example:

```{r make_toy_graph}
# let's make a small simple graph with igraph
g <- igraph::make_full_graph(5, loops = FALSE, directed = FALSE)

node_attr_names <- c("Jim", "Joe", "Bob", "Bill", "Susan")
V(g)$name <- node_attr_names

g
plot(g) # ugly default plot for sanity check
```

```{r id_to_label_test}
#g <- igraph::from_data_frame(g = node_attr_names, from =, to=)

visNetwork_g_true <- visNetwork::toVisNetworkData(g, idToLabel = TRUE)
visNetwork_g_false <- visNetwork::toVisNetworkData(g, idToLabel = FALSE)

glimpse(visNetwork_g_true$nodes)
glimpse(visNetwork_g_false$nodes)
```

Here we can see that setting that flag to `TRUE` simply copied the `id` col to a new col `label`, with identical datatype. That's ... minorly useful I guess?
