---
title: 'A2 - Adventures in function making to convert vertex names to index IDs'
author: "Andrew Sparkes"
date: "`r Sys.Date()`"
output:
  html_document:
  self_contained: true
df_print: paged
pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(echo = TRUE)

library(igraph)
library(networkdata)
library(ggplot2)
library(tidygraph)
library(ggraph)
library(visNetwork)
library(tibble)
library(purrr)
```


```{r setup_environment_to_test}
gml_filename <- "Homework_3_filtered_data.gml"

g <- wta[[40]] # storing just the 40th igraph in the wta dataset list for year 2007

g2 <- tidygraph::as_tbl_graph(g)

g3 <- g2 |>
  activate("nodes") |>
  filter(country %in% c("USA", "GBR", "AUS")) |>
  filter(!node_is_isolated())

write_graph(g3, format = "gml", file = gml_filename)

g <- igraph::read_graph(gml_filename, format = "gml")
```

This appendix focuses on tackling the problem of getting the nodes and edges from `igraph`/`ggraph` from strings (the tennis player names) to a numeric ID, so `visNetwork` can utilize the node and edge lists. I had thought this would be a straightforward task, but here I sit, hours after, but victorious. In the end, this adventure produced a single **3**-line function (that could even be one-liner'd if desired).

```{r helper_funcs_first_thoughts}
# # converts vertex `name` string to int `id` from graph `nodes`
# name_to_id <- function(name, nodes = nodes) {
#   id_df <- nodes |>
#     rownames_to_column(var = "rowname") |>
#     filter(rowname == name) |>
#     select(id)
#   id <- id_df[[1]] # extract id from df
#   id # TODO some kind of error checking. If name is not found in nodes, this func shits the bed
# }
#
#
# get_node_id_from_name <- function(nodes, name) {
#   print("get_node_id_from_name() call")
#   print(nodes)
#   print(name)
#   id_df <- nodes |>
#     filter(name == name) |>
#     select(id)
#   id_df[[1]]
# }
#
# get_node_id_from_name_vec <- Vectorize(get_node_id_from_name)
# 
# get_node_name_from_id <- function(nodes, id) {
#   print("get_node_name_from_id() call")
#   print(nodes)
#   print(id)
#   name_df <- nodes |>
#     filter(id == id) |>
#     select(name)
#   name_df[[1]]
# }
# 
# get_node_name_from_id_vec <- Vectorize(get_node_name_from_id)
#
# # test:
# get_node_id_from_name(nodes, "Lauren Albanese") # = 14
# get_node_name_from_id(nodes, 14) # "Lauren Albanese"
#
# get_node_id_from_name(nodes, "Samantha Stosur") # = 23
# get_node_name_from_id(nodes, 23) # "Samantha Stosur"
#
# get_node_id_from_name(nodes, "Alicia Molik") # = 3
# get_node_name_from_id(nodes, 3) # "Alicia Molik"

# these helper functions are fine and dandy, but because I wrote them to handle the nodes variable after converting the igraph obj to visNetwork, and in doing so, I lost much of the useful functionality of igraph. In the future, I'll keep the graph as an igraph and manipulate it then, and only then plot and do visNetwork and things... I also tried to Vectorize the functions, but that wasn't needed in the end, and produced some interesting output...

# using the igraph obj g, let's rewrite our function a bit differently
# Editor's Note: I wrote this same damn function TWICE?!?!?
# get_node_id_from_name <- function(g, name) {
#   which(V(g)$name == name)
# }
```

```{r trying_to_map_or_apply_func_to_vertices_and_failing}
# first, let's fix the from column:

# I'm gonna learn the *apply() family, the purrr package and all the various map style functions god damnit.

# it would make sense to start with vapply, as I'm wanting a numeric vector output (the id)
# vapply(edges$from, name_to_id)
# Error in vapply(edges$from, name_to_id) : argument "FUN.VALUE" is missing, with no default

# vapply(edges$from, FUN = name_to_id)
# Error in vapply(edges$from, FUN = name_to_id) :
#   argument "FUN.VALUE" is missing, with no default

# vapply(edges$from, FUN = name_to_id, FUN.VALUE = NULL)
# Error in vapply(edges$from, FUN = name_to_id, FUN.VALUE = NULL) :
#   'FUN.VALUE' must be a vector

# vapply(edges$from, FUN = name_to_id, FUN.VALUE = c(0))
# Error in rownames_to_column(nodes, var = "rowname") :
# is.data.frame(df) is not TRUE

# so it's complaining about the datatype of the data parameter I'm passing in.... Okay...
# let's investigate its data type and class and obj structure
#edges$from # dumps whole vector data
#class(edges$from) # "character" vector
#str(edges$from) # chr vector, 1:38

# So we see that our edges object is a vector (length 38) of chrs of various lengths
# class is character
# str shows vector datatype, index range/length of vector and contents

# maybe give sapply() a shot? using third parameter to pass nodes obj for get_node_id_from_name

#test <- sapply(E(g)$_from, FUN = get_node_id_from_name, nodes = nodes) # TODO WTF kind of output is that!?!
#str(test)


# Editor's note: D'oh. I spent an hour testing different ways of converting name to id, and the final function I came up with in the end below is ... the exact same as this function I wrote an hour or two ago, when I was still trying to get apply/map/etc to work. Argh!
# get_node_id_from_name <- function(name, g = g) {
#   which(V(g)$name == name)
# }
#
# sapply(E(g)$from, get_node_id_from_name, g = g) # returns list() because to and from attributes are not accessible like this with the $
```

I'm not the biggest fan of AI code generation, but perhaps I'll give it a shot and see if I can gleam anything useful. Let's throw a simple prompt at ChatGPT and see how terrible its output is...

```{r can_chatgpt_help_me_understand}
# apparently I'm struggling with igraph attributes more than I thought.

# For educational purposes, I asked ChatGPT to come up with something. Let's see a few things: 1) is it correct and working in the first place? 2) is it written well enough? and most importantly 3) can I learn something from its method of solving the problem? Let's find out...

# my prompt, verbatim:
# I have an igraph that looks like this:
# IGRAPH 81ca7e4 DNW- 29 38 -- WTA Season 2007 + attr: name (g/c), id (v/n), name (v/c), age (v/n), hand (v/c), country (v/c), surface (e/c), weight (e/n) + edges from 81ca7e4 (vertex names): [1] Julie Ditty ->Lindsay Davenport Vania King ->Meghann Shaughnessy Abigail Spears ->Ahsha Rolle [4] Abigail Spears ->Venus Williams Alexa Glatch ->Jill Craybas Alicia Molik ->Lilia Osterloh [7] Alicia Molik ->Melanie South Alicia Molik ->Serena Williams Alicia Molik ->Vania King [10] Alicia Molik ->Venus Williams Anastasia Rodionova->Alicia Molik Anastasia Rodionova->Serena Williams [13] Anastasia Rodionova->Venus Williams Anne Keothavong ->Ahsha Rolle Anne Keothavong ->Elena Baltacha [16] Ashley Harkleroad ->Meilen Tu Ashley Harkleroad ->Venus Williams Casey Dellacqua ->Nicole Pratt [19] Jamea Jackson ->Ashley Harkleroad Jamea Jackson ->Samantha Stosur Jill Craybas ->Lilia Osterloh [22] Jill Craybas ->Meilen Tu Jill Craybas ->Vania King Laura Granville ->Ashley Harkleroad + ... omitted several edges
# Note how vertices have a id attribute. Now, the edges of the igraph (the to and from columns) are actually strings (the same name attributes of the vertices), but I need to convert the to and from columns to the vertex ids. Can you write a function in R that takes a igraph object, and a name string, and renames each edge's to and from to the corresponding id, instead of the name?

# name_to_vertex_id <- function(g, name) {
#   idx <- which(V(g)$name == name)
#
#   if (length(idx) == 0) {
#     stop(paste("Vertex name not found:", name))
#   }
#
#   V(g)$id[idx]
# }
#
# convert_edge_names_to_ids <- function(g) {
#   if (is.null(E(g)$from) || is.null(E(g)$to)) {
#     stop("Edges must have 'from' and 'to' attributes containing vertex names")
#   }
#
#   E(g)$from_id <- vapply(
#     E(g)$from,
#     function(x) name_to_vertex_id(g, x),
#     numeric(1)
#   )
#
#   E(g)$to_id <- vapply(
#     E(g)$to,
#     function(x) name_to_vertex_id(g, x),
#     numeric(1)
#   )
#
#   return(g)
# }
#
# g <- convert_edge_names_to_ids(g) # surprise surprise, this line throws:
# # Error in convert_edge_names_to_ids(g) :
# #   Edges must have 'from' and 'to' attributes containing vertex names

# all that relatively nicely formatted, proper /looking/ code, and it doesn't do jackshit. To be expected when AI slops tries to problem solve. It /looks/ right, but it's always a crap shoot if it actually functions...
```

Well, that detour provided very little benefit. First off, the ChatGPT code didn't actually function in the first place. It's cleanly formatted and clear to read, but that's rather a moot point if it doesn't actually get the job done... And as for learning anything from it? It does include some basic error checking (null checks, and not found check in the helper function), but nothing that I didn't already know unfortunately. Assuming its calling `vapply` correctly, I'm almost on the right track for the `*apply` family of functions.

Now, let's shake off all that above crap, and try something from a slightly different approach:

```{r different_approach}
# a redo at the same function, takes in igraph g, string name, returns that vertex's internal id

# node_name_to_id <- function(g, name) {
#   result <- g.vs.find(name = name).index
#
#   result
#   }
#
# # a quick test:
# node_name_to_id(g, "Lindsay Davenport")
# resulted in:
# Error: unexpected symbol in:
# "node_name_to_id <- function(g, name) {
#   g.vs.find(name = name).index"

# symbol... okay... I don't see any special symbols in there, save for possible the . accessor for index? Ah, I'm dumb. I see. After more digging in igraph documentation, I come to this modification, with the $:

# node_name_to_id <- function(g, name) {
#   result <- g.vs.find(name = name)
#
#   result$index
# }
# node_name_to_id(g, "Lindsay Davenport")
# results in:
# # Error in g.vs.find(name = name) : could not find function "g.vs.find"

# it would appear that the .find() method of the VertexSequence obj exists in the Python igraph library, but not in the R package ... hmmm... apparently I wasn't careful enough when binging through igraph documentation

# node_name_to_id <- function(g, name) {
#   result <- V(g)$name == name # should return the vertex itself
#
#   result$index
# }
# node_name_to_id(g, "Lindsay Davenport")
# outputs:
# Error in result$index : $ operator is invalid for atomic vectors

# fortunately, that gives me a lead as to where the issue is somewhat. I got many different datatypes flying around and being coerced and such

# node_name_to_id <- function(g, name) {
#   result <- V(g)[name == name] # should return the vertex itself? But it appears to barf out the full igraph.vs VertexSequence obj... WTF
#
#   result
# }
# test <- node_name_to_id(g, "Lindsay Davenport")
# class(test)
# str(test)
# spits out:
# [1] "igraph.vs"
#  'igraph.vs' Named int [1:29] 1 2 3 4 5 6 7 8 9 10 ...
#  - attr(*, "names")= chr [1:29] "Abigail Spears" "Ahsha Rolle" "Alexa Glatch" "Alicia Molik" ...
#  - attr(*, "env")=<weakref>
#  - attr(*, "graph")= chr "81ca7e4c-0627-11f1-bb26-e55ca1e4ce34"


node_name_to_id <- function(g, name) {
  result <- which(V(g)$name == name) # should return the vertex index itself?

  result
}
test <- node_name_to_id(g, "Lindsay Davenport") # should be 17
test # [1] 17
class(test) # [1] "integer"
str(test) #  int 17

# well, now we're talking... That ended up simpler than I was anticipating... I feel a bit dumb, but also feel a bit more learned, so I'll take it.
```


```{r final_function_cleanup_part_one}
# one more cleanup/formatting pass to make it pretty, functional and concise (just like that 3 line version of it, which is identical, a few hours ago up above, d'oh):
igraph_node_name_to_id <- function(g, name) {
  which(V(g)$name == name)
}

# one final test to ensure the function works as intended
test <- igraph_node_name_to_id(g, "Samantha Stosur") # should be 24
test # [1] 24
class(test) # [1] "integer"
str(test) #  int 24
```

That looks like it checks out to me! We have a working `igraph_node_name_to_id(g, name)`!

Unfortunately, that's only part of the functionality we need. We also need to make a function `igraph_convert_names_to_ids(g)` that will change the to/from edge attributes to integer IDs based on the index/ID of the vertices.

```{r set_back_up_environment_for_testing}
g <- igraph::read_graph(gml_filename, format = "gml")
```

```{r finish_that_last_crucial_function}
#igraph_convert_names_to_ids <- function(g) {
  # node_name_to_id(g, "Samantha Stosur") => 24

  v_attr <- vertex_attr(g)
  lookup_table <- (v_attr$id) # id starts at zero, may need to start at 1? maybe?
  names(lookup_table) <- v_attr$name

  # now, lookup_table["name"] should return the id

  # now, just need to do a map/apply over the vector of edge$from values, edge$to values, and then doublecheck they match the node ids

  #set_edge_attr(graph = g, name = "from", value = lookup_table[E(g)$from]
#}

# well, if I can't seem to figure out how to replace the strings with their integer ids, one method is to remove and readd the edges, but that doesn't seem great. Another idea is to export the node and edge information out, manipulate it, then make a new igraph object and pass it to visNetwork
```


```{r maybe_vector_func}
node_name_to_id_vec <- Vectorize(node_name_to_id)

# Example data frame with string "from" and "to" columns
edge_list_ints <- data.frame(
  from = node_name_to_id_vec(E(g)$from),
  to = node_name_to_id_vec(E(g)$to)
)

g <- graph_from_data_frame(edge_list_ints, directed = TRUE)

g
igraph_convert_names_to_ids(g)
```

```{r convert_igraph_to_visnetwork}
data <- visNetwork::toVisNetworkData(g, idToLabel = FALSE)

nodes <- data$nodes
edges <- data$edges
```


```{r final_function_cleanup_part_2}
```


