---
title: "STAT332 Statistical Analysis of Networks Sec 01 - Homework 4 - Prof. Joe Reid"
author: "Andrew Sparkes"
date: "`r Sys.Date()`"
due: "06-02-2026"
output:
  html_document:
    self_contained: true
    df_print: paged
  pdf_document: default
link-citations: true
natbib: true
csl: ieee.csl
bibliography: NetworkReferences.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(igraph)
library(knitr)
library(gt)
library(data.table)
library(tidygraph)

source("../graph.R")
```

For this lab, please refer to the facebook dataset. This dataset is from [@rozemberczki2019multiscale].

You may use the following code to read and generate the graph.

```{r read_csv_files}
nodes <- read.csv("musae_facebook_target.csv", header = TRUE, as.is = TRUE)
links <- read.csv("musae_facebook_edges.csv", header = TRUE, as.is = TRUE)

```


```{r make_graph}
g <- igraph::graph_from_data_frame(d = links, vertices = nodes, directed = FALSE)

```

0.) Your first step, before doing any other work, will be to create a subgraph that is limited to only those nodes where the `page_type` is "tvshow". Then overwrite this network with an induced subgraph where the isolates are removed (you should have **3,282** nodes)

```{r create_subgraph}
V(g)$degree_total <- degree(g, mode = "all", loops = FALSE) # , normalized = TRUE removed so it's more human readable

g <- g %>% 
  as_tbl_graph() %>%
  activate("nodes") %>%
  filter(page_type == "tvshow" & degree_total > 0)

cat(paste0(
  "Number of nodes: ",
  length(V(g))
))
```

I wouldn't recommend trying to plot the network right away. Start by looking at the size of the data use in the environment tab. Please write the code to produce an answer to the following questions and then interpret the results when asked.

1.) Produce a summary of the graph object and answer the following by producing a nice summary table of the following information using the work from the lab.

```{r summarize_the_graph}
# barf out the easy stuff first
summary(g)
```

```{r make_pretty_table}
# now, let's calculate some better measures and spit them out in a nicely formatted gt() table:

print_formatted_graph_summary_table(g)
```

Editor's note: why is there something that is flagging the network as having loops? That doesn't make sense... Let's remove that...

```{r}
g <- simplify(g)
```

And to confirm:
```{r}
print_formatted_graph_summary_table(g)
```

Wait, what? Simplify is supposed to removed multiple edges and any loops present and yet this... Hmmmm....

# TODO Fix issue here?

2.)  Produce the **5** measures of graph structure discussed in the lesson from this week. Report these in a table that includes the measure name, the value of the statistic, and an interpretation of the statistic.

```{r create_basic_measures_table}
output_five_basic_measures_table <- function(g) {
  tbl <- data.frame(
    measure_name = c(
      "asdf",
      "",
      "",
      "",
      ""
    ),
    measure_value = c(
      "",
      "",
      "",
      "",
      ""
    )
  )

  tbl |> gt()
}

output_five_basic_measures_table(g)
```

3.) Create a table of the interconnectedness measures for this graph and provide an interpretation of this results.

```{r create_interconnectedness_measures_table}
output_interconnectedness_measures_table <- function(g) {
  tbl <- data.frame(
    measure_name = c(
      "Edge Density",
      "Cohesion",
      "Global Clustering",
      "Assortativity"
    ),
    measure_value = c( # TODO placeholder
      1:4 
    )
  )

  tbl |> gt()
}

output_interconnectedness_measures_table(g)
```

**Answer:**
TODO interpretation

4.) Create a table with the degree, closeness, betweenness and eigenvalue centralization measures for this graph and interpret the results.

```{r create_more_measures_table}
output_more_measures_table <- function(g) {
  tbl <- data.frame(
    measure_name = c(
      "Degree",
      "Closeness",
      "Betweenness",
      "Eigenvalue"
    ),
    measure_value = c( # TODO placeholder
      1:4
    )
  )

  tbl |> gt()
}

output_more_measures_table(g)
```

**Answer:**
TODO interpretation

5.) Produce a table of the efficiency measures of this graph and provide an interpretation of the results.

```{r create_efficiency_measures_table}
output_efficiency_measures_table <- function(g) {
  tbl <- data.frame(
    measure_name = c(
      "Average Shortest Path",
      "Global Efficiency",
      "Average Local Efficiency"
    ),
    measure_value = c( # TODO placeholder
      1:3
    )
  )

  tbl |> gt()
}

output_efficiency_measures_table(g)
```

**Answer:**
TODO interpretation

6.) Given the measures that you have produced in this assignment, would you consider this network to be centralized or de-centralized? Similarly, would you consider this network to be highly interconnected or have low interconnectedness? Based on these results, what do you think the advantages and disadvantages of this network would be in terms of robustness and resilience. What are the critical strengths and weaknesses of this network?

**Answer:**
# TODO interpretation

## References
