---
title: "STAT332 Statistical Analysis of Networks Sec 01 - Project 2 - Prof. Joe Reid"
author: "Andrew Sparkes"
date: "`r Sys.Date()`"
due: "13-03-2026"
output:
  html_document:
    self_contained: true
    df_print: paged
    code_folding: show
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(igraph) 
library(igraphdata)

library(data.table)
library(gt)

library(patchwork)
library(ggraph)
library(visNetwork)

source("../graph.R")
```

```{r useful_functions}
# takes in a string, returns a span with the color attribute set to whatever the color parameter is (anything that is accepted by HTML, so RGB hex codes, named colors, etc)
textcol <- function(text, color){
  paste0('<span style="color: ',
         color,
         ';">',
         text,
         '</span>'
         )
}

```
---

# Grand Overview of the Project 

## Instructions

For your final project, you will conduct a comprehensive network analysis of a network of your choosing. Your goal is not merely to compute metrics, but to tell a coherent and analytical story about something important, critical, or revealing about your network.

You should think of this project as a case study in network science:

 * What kind of system is this?
 * What matters structurally?
 * What does the network reveal that would be hard to see otherwise?

You are encouraged to choose a network that connects to your interests (social, biological, technological, informational, political, cultural, etc.).

## 0.) Project Requirements at a Glance

Your project must include all of the following components:

| Key Components --------| More Key Components---------------------------------- |
| :--------------------: | :---------------------------------------------------: |
|Scale 	                 | Required Topics                                       |
|Visualization 	         | Network layout(s) and visual interpretation           |
|Macro-scale 	           | Global structure & overall network properties         |  
|Meso-scale 	           | Communities, structure, and group-level patterns      | 
|Micro-scale 	           | Node and/or edge-level importance                     |
|Dynamics / robustness 	 | Network attack or failure analysis                    | 
|Interpretation 	       | A unifying narrative or “story”                       |

How you connect these pieces—and which specific measures you choose—is up to you.


---

## 1. Network Description & Research Question

**Goal**: Introduce the network and motivate your analysis.

Include:

 * What does the network represent?\
 * What do nodes and edges mean?\
 * Is it directed/undirected?\
 * Weighted/unweighted?\
 * Where did the data come from?\

Then clearly state _**one** guiding question_, such as:\
 * “Which actors are most critical to the functioning of this system?”\
 * “How robust is this network to targeted disruption?”\
 * “Does influence concentrate within or across communities?”\
 * “What structural roles explain observed outcomes?”\

This question should guide _all_ later analysis:


---

## 2. Network Visualization & Exploratory Structure

**Goal**: Use visualization to reveal structure, not just aesthetics.

**Required**:

  * **At least two different visualizations** (one will be talked about later in the instructions)
  * Thoughtful choices about:
    * Layout algorithms
    * Node size and/or color
    * Edge width or opacity (_if_ weighted)
    * Labels
    * Legends

**Your task is to explain**:

  * What patterns are visible?\
  * Are there clusters, hubs, bridges, or isolates?\
  * What does visualization suggest before formal analysis?\
    
**Answer:**

# TODO ME! basic interpretation. clusters, hubs, isolates, and interesting patterns, bridges, etc


---

## 3. Macro-Scale Structure

**Goal**: Characterize the network as a whole.

You _must_ analyze **at least three (3) macro-scale properties**, that we covered in class and apply to your particular network.

Explain what these metrics say about your network in the context of your question.  This might include topics such as:

 * Connectivity
 * Reachability
 * Overall organization

# TODO explain how these ties lead back into your guiding question


---

## 4. Meso-Scale Structure: Communities & Groups

**Goal**: Identify and interpret intermediate structure.

**Required**:

 * Apply at least one community detection or structural method\
 * Assign nodes to communities or groups\
 * Visualize the result (this is **your second required visualization**)\

Then, interpret questions like the following (pertaining to your question):

 * Are communities meaningful in the real-world context?\
 * Are they densely connected or loosely affiliated and what does that mean?\

This section should connect naturally to later discussions of brokerage and robustness.


---

## 5. Micro-Scale Analysis: Nodes and/or Edges

**Goal**: Identify important actors or relationships.

You _must_:

 * *_Select_* *at least two* node-level measures and/or one edge-level measure
 * *_Justify_* _why_ these measures are appropriate
 * *_Identify_* the most extreme or influential nodes/edges in the context of the question you are addressing

Possible (not required) themes:

 * Centrality
 * Brokerage or mediation
 * Prestige or influence
 * Embeddedness
 * Constraint or redundancy
 * Critical ties

*Do not just report rankings—interpret why these entities matter!*


---

## 6. Network Robustness & Attack Analysis

**Goal**: Evaluate fragility and resilience.

You must perform **at least _one_ (1) network attack or failure analysis**, such as:

 * Targeted removal based on a node or edge metric
 * Random failure
 * Comparison of different attack strategies

This may be a full simulation or it may be an analysis of single stage attacks on critical structures. Remember, attacks can include infection, removal, addition, or modulation of components of the graph and may be malicious, random, or even exploratory what-if types of analysis. Get creative!

Analyze how actions affect some global and local level measure of your choice that reflect your key topic or theme.

Things you may interpret:

 * What does this tell us about vulnerability?
 * Which structural features contribute to robustness or fragility?


---

## 7. Synthesis: Telling the Network’s Story

**Goal**: Bring everything together.

In this final section, _explicitly answer_:

 * *What* is the key insight your network analysis revealed?
 * *How* do macro-, meso-, and micro-scale findings connect?
 * *What* would someone misunderstand if they didn’t use network analysis?


---

# Deliverables

 * A written report (PDF or HTML), approximately *10–15* pages
 * Clear *section headings* _aligned with the structure_ above
 * Figures must be *legible*, *labeled*, and *discussed in text*
 * *Code must be included in a separate file for reproducibility*.
 * *Network/graph/dataset must also be included in a separate file*.
 * Put *all* files into a `.zip` folder and submit that.

*Evaluation Criteria* (What I Will Be Looking For)\
You will be graded primarily on:

 * *Appropriateness* of methodological *choices*
 * *Quality of interpretation*
 * *Integration* across scales
 * *Clarity* and *coherence* of your *narrative*
 * *Depth of insight*, not number of metrics

There is no single "correct" answer. There are well-argued and poorly-argued analyses.






































```{r load_dataset}
# gml_file <- "as-22july06.gml"
# gml_file <- "lesmis.gml"
rds_file <- "got.rds"

# g <- igraph::read_graph(gml_file, format = "gml")
g <- readr::read_rds(rds_file)
```

```{r}
g
```

```{r plot_the_graph_data}
plottt(g)
```

```{r}
ID = c("Order", "Size", "Connected", "Directed", "Acyclic", "Weighted", "Simple", "Has Loops", "Is Multigraph", "Bipartite", "Tree", "Forest")
  
Value = c(gorder(g), ecount(g), is_connected(g), is_directed(g), is_acyclic(g), is_weighted(g), is_simple(g), any_loop(g), any_multiple(g), is_bipartite(g), is_tree(g), is_forest(g))
  
table <- data.table(ID, Value)
table %>% gt() %>% fmt_tf(columns = Value, tf_style = "yes-no")
```

```{r}
diameter(g)
```

```{r}
get_diameter(g)
```

```{r}
diameter(g, weights = 1 / E(g)$weight)
```

```{r}
get_diameter(g, weights = 1 / E(g)$weight)
```


```{r}
cat("There are",
    length(bridges(g)),
    "bridges in this graph, and there are",
    length(articulation_points(g)),
    "articulation points in this graph."
    )

```

```{r}
bridges(g)
bridges_all <- bridges(g)

bridges_all
```


```{r}
E(g)$from


# E(g)$is_bridge <- E(g)[which(E(g)$from %in% bridges(g) | which((E(g)$to %in% bridges(g)))
```

```{r}
articulation_points(g)
```

```{r}
farthest_vertices(g)
```

```{r}
edge_density(g, loops = FALSE)
```

```{r}
cohesion(g)
```

```{r}
vertex_connectivity(g)
```

```{r}
reachability_cohesion(g)
```

```{r}
transitivity(g, type = "globalundirected", isolates = "NaN") # TODO ASK what is isolates="NaN" here and not just NA?
```

```{r}
global_weighted_transitivity(g, mode = "undirected")
```

```{r}
assortativity_degree(g)
```

```{r}
# local_neighborhood <- igraph::make_neighborhood_graph(
#   g,
#   order = 4,
#   nodes = V(g),
#   mode = "all",
#   mindist = 0
# )
# plot(local_neighborhood)
```

```{r}
# ego_graph <- igraph::make_ego_graph(
#   g,
#   order = 4,
#   nodes = V(g),
#   mode = "all",
#   mindist = 0
# )
# plot(ego_graph)
```



