---
title: "STAT332 Statistical Analysis of Networks Sec 01 - Homework 6 - Prof. Joe Reid"
author: "Andrew Sparkes"
date: "`r Sys.Date()`"
due: "30-01-2026"
link-citations: true
natbib: true
output:
  html_document:
    self_contained: true
    df_print: paged
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r user_configurable_settings}
```

## Background

You are given an **undirected, weighted network** representing
interactions between characters in *Game of Thrones*. Nodes represent
characters, and edge weights represent the **strength of interaction**
between pairs of characters.

You should treat this as a **social network** and analyze it using the
concepts and tools from class.

**Important:**\
For each task below, **you must decide which network measure(s) are
appropriate**. There is no single correct answer—but there *are*
incorrect or poorly justified ones.

For every task:

1.  State which measure(s) you used

2.  Briefly explain *why* that measure fits the goal

3.  Identify the character(s) with the highest values

4.  Include a small table or list of the top results

# Prepping the dataset
```{r load_dataset}
# TODO load the GoT dataset

```


## Task 1: Central Characters

**Goal:**\
Identify the character(s) who are most *central* to the overall network
structure.

**Guidance:**\
“Central” here refers to characters who are structurally important
across the network, not just locally popular. You should consider how
information, influence, or interaction potential flows through the
network.

**Deliverables:**

- The measure(s) you selected

- The top 3 characters by that measure

- A brief justification of your choice and what it indicates about these
  characters.

**Answer**:
For determining which nodes are structurally important _across_ the network, from our centralization metrics, I chose to calculate XYZ

Determining the actual value for this dataset:
```{r}
# TODO calculate

```


## Task 2: Powerful Intermediaries

**Goal:**\
Identify characters who act as **intermediaries** between otherwise
weakly connected parts of the network.

**Guidance:**\
These characters may not have the strongest direct ties, but their
position allows them to connect groups or clusters. Focus on *structural
mediation*, not popularity.

**Deliverables:**

- The measure(s) you used to capture this information

- The top 3 characters and what kind of connection they provide

- A short explanation of how your measure captures this information

**Answer**:
For determining which nodes are acting as ties between groups/clusters, I will determine any bridge points and measure XYZ

Determining the actual values for this dataset:
```{r}
# TODO calculate

```

The top three characters, ranked by their ability and effectiveness at linking clusters together:

## Task 3: Prominence

**Goal:**\
Identify the most **prestigious** characters in the network.

**Guidance:**\
Prestige is not just about how many connections a character has, but
*who* they are connected to and how strong those ties are. Think
carefully about how prestige differs from simple centrality.

**Deliverables:**

- The prestige-related measure(s) you selected

- The top 3 characters by prestige

- A brief justification of your choice and what it means for these
  character roles.

**Answer**:
For determining which nodes/characters are the most prestigious/how strong the adjacent edges are, I chose XYZ

Determining the actual value for this dataset:
```{r}
# TODO calculate

```

The top three characters by prestige are:
```{r top_three_prestigious_characters}

```


## Task 4: Social Embeddedness

**Goal:**\
Identify characters who are most **embedded in tightly knit social
environments**.

**Guidance:**\
These characters tend to be surrounded by others who are also connected
to one another. They may be influential within a clique or local group,
even if they are not globally central.

**Deliverables:**

- The measure(s) you used

- The top 3 characters by this criterion

- A brief explanation of what “embeddedness” means in network terms

**Answer**:
For determining which nodes/characters are the most asdfsdf, I chose XYZ

Determining the actual value for this dataset:
```{r}
# TODO calculate

```

The top three characters by prestige are:
```{r top_three_prestigious_characters}

```

"Embededness" means alskdfjlasdkf

## Task 5: Strength of Influence

**Goal:**\
Identify characters whose **relationships are strongest overall**, not
merely numerous.

**Guidance:**\
This task explicitly requires you to use the **edge weights**. Focus on
characters whose total interaction strength is high, and reflect on how
this differs from unweighted analyses.

**Deliverables:**

- The measure(s) you used

- The top 3 characters by strength-based influence and what it means for
  these characters.

- A brief explanation of why weights matter for this task

**Answer**:
For determining which nodes/characters have the strongest relationships overall, I chose XYZ

Determining the actual value for this dataset:
```{r}
# TODO calculate

```

The top relationships, by weight/strength, are:
```{r top_relationships}

```

Weights matter in this scenario because alksdjflaksdjflasdjflksd

## References
