---
title: "STAT332 Statistical Analysis of Networks Sec 01 - Homework 6 - Prof. Joe Reid"
author: "Andrew Sparkes"
date: "`r Sys.Date()`"
due: "27-02-2026"
link-citations: true
natbib: true
output:
  html_document:
    self_contained: true
    df_print: paged
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(igraph) # Our standard graph object.
library(readr) # Reading in the data
library(gbRd)
library(gt)
library(ggraph) # Plotting
library(tidygraph) # Managing our graph object
library(RColorBrewer) # Getting Colors
library(scales) # scaling factors
library(grid) # for unit()
library(ggrepel) # for stronger label repel
library(visNetwork) # for network graphs
library(ggplot2) # for ploots
library(igraphdata)

source("../graph.R")
```

```{r user_configurable_settings}
got_rds_file <- "got.rds"
```

## Background

You are given an **undirected, weighted network** representing interactions between characters in *Game of Thrones*. Nodes represent characters, and edge weights represent the **strength of interaction** between pairs of characters.

You should treat this as a **social network** and analyze it using the concepts and tools from class.

**Important:**\
For each task below, **you must decide which network measure(s) are appropriate**. There is no single correct answer—but there *are* incorrect or poorly justified ones.

For every task:

1.  State which measure(s) you used

2.  Briefly explain *why* that measure fits the goal

3.  Identify the character(s) with the highest values

4.  Include a small table or list of the top results

---

# Prepping the dataset
```{r load_dataset}
g <- readr::read_rds(got_rds_file)
```

## Task 1: Central Characters

### Goal

Identify the character(s) who are most *central* to the overall network structure.

### Guidance

- The measure(s) you selected

- The top 3 characters by that measure

- A brief justification of your choice and what it indicates about these
  characters.

**Answer**:
For determining which nodes are structurally important _across_ the network, from our centralization metrics, I chose to calculate degree centrality for each node, with in degree centrality indicating more of a popular "central" and out degree centrality indicating more of an authority kind of central character.

The top three characters, ranked by their central-ness (degree) to the overall network structure:

```{r calculate_all_the_degrees}
V(g)$deg_total = degree(g, v = V(g), mode = "all", loops = FALSE, normalized = TRUE)
V(g)$deg_in = degree(g, v = V(g), mode = "in", loops = FALSE, normalized = TRUE)
V(g)$deg_out = degree(g, v = V(g), mode = "out", loops = FALSE, normalized = TRUE)
```


```{r total_degree_top_three}
deg_total_order <- order(V(g)$deg_total, decreasing = TRUE)
deg_total_sorted_vertices <- V(g)[deg_total_order]
deg_total_sorted_vertices[1:3]
```

```{r in_degree_top_three}
deg_in_order <- order(V(g)$deg_in, decreasing = TRUE)
deg_in_sorted_vertices <- V(g)[deg_in_order]
deg_in_sorted_vertices[1:3]
```

```{r out_degree_top_three}
deg_out_order <- order(V(g)$deg_out, decreasing = TRUE)
deg_out_sorted_vertices <- V(g)[deg_out_order]
deg_out_sorted_vertices[1:3]
```

**Answer:**\
The top three characters by degree centrality (mode = "all") are Tyrion Lannister, Jon Snow and Sansa Stark. As a watcher of the show, this makes complete sense.

---

## Task 2: Powerful Intermediaries

**Goal:**\
Identify characters who act as **intermediaries** between otherwise
weakly connected parts of the network.

**Guidance:**\
These characters may not have the strongest direct ties, but their
position allows them to connect groups or clusters. Focus on *structural
mediation*, not popularity.

**Deliverables:**

-    The measure(s) you used to capture this information

-    The top 3 characters and what kind of connection they provide

-    A short explanation of how your measure captures this information

**Answer**:

# TODO For determining which nodes are acting as ties between groups/clusters, I will determine any bridge points and measure XYZ

Determining the actual values for this dataset:
```{r}
# TODO calculate

```

# TODO The top three characters, ranked by their ability and effectiveness at linking clusters together:

---

## Task 3: Prominence

**Goal:**\
Identify the most **prestigious** characters in the network.

**Guidance:**\
Prestige is not just about how many connections a character has, but
*who* they are connected to and how strong those ties are. Think
carefully about how prestige differs from simple centrality.

**Deliverables:**
  
-   The prestige-related measure(s) you selected

-   The top 3 characters by prestige

-   A brief justification of your choice and what it means for these character roles.

**Answer**:

# TODO For determining which nodes/characters are the most prestigious/how strong the adjacent edges are, I chose XYZ

Determining the actual value for this dataset:
```{r}
# TODO calculate

```

# TODO The top three characters by prestige are:
```{r top_three_prestigious_characters}

```

---

## Task 4: Social Embeddedness

**Goal:**\
Identify characters who are most **embedded in tightly knit social environments**.

**Guidance:**\
These characters tend to be surrounded by others who are also connected to one another. They may be influential within a clique or local group, even if they are not globally central.

**Deliverables:**
  
-   The measure(s) you used

-   The top 3 characters by this criterion

-   A brief explanation of what “embeddedness” means in network terms\

**Answer**:
# TODO For determining which nodes/characters are the most asdfsdf, I chose XYZ

# TODO The top three characters ranked by social embededness are:

```{r}
# TODO calculate

```

# TODO "Embededness" means alskdfjlasdkf

---

## Task 5: Strength of Influence

**Goal:**\
Identify characters whose **relationships are strongest overall**, not merely numerous.

**Guidance:**\
This task explicitly requires you to use the **edge weights**. Focus on characters whose total interaction strength is high, and reflect on how this differs from unweighted analyses.

**Deliverables:**

-   The measure(s) you used

-   The top 3 characters by strength-based influence and what it means
for these characters.

-   A brief explanation of why weights matter for this task

**Answer**:

# TODO For determining which nodes/characters have the strongest relationships overall, I chose XYZ

Determining the actual value for this dataset:
```{r}
# TODO calculate

```

The top relationships, by weight/strength, are:
```{r top_relationships}
# TODO top relationships
```

# TODO Weights matter in this scenario because alksdjflaksdjflasdjflksd
