---
title: "STAT332 Statistical Analysis of Networks Sec 01 - Lab 3 - Prof. Joe Reid"
author: "Andrew Sparkes"
date: "`r Sys.Date()`"
due: "30-01-2026"
output:
  html_document:
    self_contained: true
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(gt) # for fancy tables
library(data.table) # data tables
library(igraph) # Our standard graph object.
library(readr) # Reading in the data
library(ggraph) # Plotting
library(tidygraph) # Managing our graph object
library(RColorBrewer) # Getting Colors
library(scales) # scaling factors
library(grid) # for unit()
library(ggrepel) # for stronger label repel
library(visNetwork) # for network graphs
library(ggplot2) # for ploots

source("../graph.R")
```

```{r user_configurable_settings}
# for first part:
media_network_filename <- "Media Network.rds"

# for later part:
actor_roles_filename <- "Actor Roles.csv"
table_2_filename <- "2006 Koschade Table 2.csv"
```

## 1.) Setup

Welcome to Lab 3. This week we are going to be working with alternatives to `igraph`. This means that we want to get familiar with the tools from `ggraph` (<https://ggraph.data-imaginist.com/index.html>) and `visNetwork` (<https://datastorm-open.github.io/visNetwork/>).

We're going to be using the same data as lab 2 and trying to reproduce the same graph as last time, so lets get started.

## 2.) Lab Walkthrough

### 2.a) Getting started.

Using `ctrl-alt-i` you can create a section where you are going to embed R code. We're going to start by loading the raw data file:

```{r read_data_file}
g <- read_rds(media_network_filename)
```

Just to remind ourselves of what we're working with, let's make sure we print out a quick summary, and dumping all the details because it's such a small network:

```{r graph_summary}
g
print_igraph_attr(g)
```

### 2b.) Adding degree strength as attributes

Just like last week, we want to add a few more attributes.

```{r set_vertex_strength_attrs}
g <- set_vertex_attr(g, name = "strength_in", value = strength(g, mode = "in"))
g <- set_vertex_attr(g, name = "strength_out", value = strength(g, mode = "out"))
g <- set_vertex_attr(g, name = "strength_total", value = strength(g, mode = "total"))
```

Let's make sure we have our new attributes.

```{r summarize_again}
summary(g)
```

### 2d.) Let's graph

This time, we're going to use

```{r plot_the_graph}
set.seed(1)
g |> # Start with the graph object
  ggraph(layout = "lgl") + # initializes the plot
  geom_node_point() + # adds the nodes to the plot
  geom_edge_link() # adds the edges to the plot
```

Alright, that's the basics.

```{r make_more_graph}
# For reference:
# IGRAPH 52040ab DNW- 17 49 --
# + attr: name (v/c), media (v/c), media.type (v/n), type.label (v/c), audience.size (v/n), strength_in (v/n),
# | strength_out (v/n), strength_total (v/n), type (e/c), weight (e/n)

set.seed(1)
p <- g |> # Start with the graph object
  ggraph(layout = "lgl") +
  geom_node_point(
    aes(
      size = audience.size,
      fill = type.label,
      shape = type.label
      ),
    color = "grey30",
    alpha = 0.5,
    stroke = 0.4
  ) +
  geom_node_text(
    aes(label = media),
    repel = TRUE,
    size = 3
  ) +
  geom_edge_link(
    aes(
      color = type,
      edge_linetype = type,
      width = weight/10
      ),
    alpha = 0.5,
    lineend = "round",
    arrow = arrow(length = unit(3, "mm"), type = "closed"),
    end_cap = circle()
  ) +

  labs(
    title    = "Media Networks Link to and Reference Each Other",
    subtitle = "A network plot of networks.",
    caption  = "Data from: https://kateto.net/network-visualization."
  )

p
```

Okay, it's not perfect. We can do better, but the format is just so much better. Let's fix the scales next:

```{r fixing_scales}
# Shape mapping (3 shapes)
node_types <- sort(unique(V(g)$type.label))
shape_vals <- c(21, 22, 24)[seq_along(node_types)]
shape_map <- setNames(shape_vals, node_types)

set.seed(1)
p <- ggraph(g, layout = "lgl") +
  geom_node_point(
    aes(
      size = audience.size,
      fill = type.label,
      shape = type.label
      ),
    alpha = 0.6
  ) +
  geom_node_text(
    aes(
      label = media
      ),
    repel = TRUE,
    size = 3
  ) +

  geom_edge_arc(
    aes(
      color = type,
      linetype = type, 
      width = weight
      ),
    alpha = 0.4,
    lineend = "round",
    arrow = arrow(length = unit(3, "mm"), type = "closed"),
    end_cap = circle(2, "mm"),
    strength = 0.3
  ) +


  scale_size_continuous(name = "Audience Size", range = c(1, 8)) +
  scale_shape_manual(values = shape_map, guide = "none") +
  scale_edge_width(range = c(0.5, 1.5), guide = "none") +
  
  guides(
    size = guide_legend(nrow = 2, byrow = TRUE),
    fill = guide_legend(
      nrow = 1, byrow = TRUE, title = "Vertex Color/Shape",
      override.aes = list(shape = shape_vals)
    ),
    edge_color = guide_legend(
      ncol = 2, byrow = TRUE,
      title = "Reference Type"
    )
  ) +

  labs(
    title    = "Media Networks Link to and Reference Each Other",
    subtitle = "A network plot of networks.",
    caption  = "Data from: https://kateto.net/network-visualization."
  )

p
```

If you're paying attention, you might notice that I stored the graph as the variable `p` and then called it by just using `p` at the end of the code cell. I did this for a reason. In `ggplot2`, we can theme our graphs. <https://ggplot2.tidyverse.org/reference/theme.html>

```{r adding_theme}
p + theme(
  plot.title = element_text(face = "bold", size = 14),
  legend.box.background = element_rect(color = "black", size = 0.5),
  legend.box.margin = margin(1, 1, 1, 1),
)
```

**Editor's Note**: Before I forget, I had to modify a couple of the chunks above, commenting out a few parameters related to edge line type. I'll have to look into what is the source of that issue.


## 3.) Your turn:

In the data for this lab, there are two files (and a paper describing where they came from). The "Actor Roles" file provides the list of nodes and their corresponding attributes. The "Table 2" file provides a matrix of associations and their strengths that you can read in as a matrix and extract the edge list from. Please create a graph which uses shapes, colors, and edge-widths to represent this network. Also, please make sure to use both color and type on edges to denote any edge relationships between Command Team members and any other members in the network. Use `ggraph` to produce this plot. Be certain to add titles, subtitles, and footnotes/captions as appropriate and include a corresponding legend.

```{r read_in_data_files}
actor_roles <- readr::read_csv(actor_roles_filename, col_names = TRUE, progress = TRUE)
table_2 <- readr::read_csv(table_2_filename, col_names = TRUE, progress = TRUE)
```

```{r convert_to_igraphs}
# actor_roles appears to map each actor to a number of roles, both strings
# table 2 seems to represent an adjacency matrix, with weights
# will need to merge data from both into graph obj

df <- as.data.frame(table_2) # first, convert to df
rownames(df) <- df$...1 # name the rows = the first col (...1) which are the terrorist names
df <- df |> select(-...1) # select all cols BUT that first col (...1)
adj_matrix <- as.matrix(df) # convert to matrix so igraph won't complain

# finally, make the igraph obj
g_actors <- igraph::graph_from_adjacency_matrix(adj_matrix, mode = "directed", weighted = TRUE, diag = FALSE)

# utilizing the data from the actors/role csv, assign the roles to each vertex as an attr
g_actors <- set_vertex_attr(g_actors, name = "Role", value = actor_roles$Role) # TODO double check these are assigned in the proper order

g_actors
print_igraph_attr(g_actors)
```

```{r identify_command_team_relations}
command_team_edges <- E(g_actors)[.inc(V(g_actors)[Role == "Command Team"])] # see https://igraph.org/r/html/1.2.3/igraph-es-indexing.html

# We'll accentuate the command team themselves with a colored node, as well coloring any contact with the team members
# note: we're just setting flags here so ggraph can color during plotting
V(g_actors)$command_team <- FALSE
V(g_actors)$command_team <- V(g_actors)$Role == "Command Team" # probably better ways

E(g_actors)$command_team <- FALSE
E(g_actors)[.inc(V(g_actors)[Role == "Command Team"])]$command_team <- TRUE

# note: command team members: id 1 Muklas, 4 Samudra, 6 Idris (should be 62 edges)
```

```{r setup_the_ploot}
command_colors <- c("#ff0000", "#000000")
V(g_actors)$str_in <- strength(g_actors, mode = "in", weights = E(g)$weights)

# create graph using ggraph
set.seed(1)
p <- ggraph(g_actors, layout = "lgl") +
  geom_edge_arc(
    aes(
      color = command_team
      ),
    alpha = 0.5,
    lineend = "round",
    arrow = arrow(length = unit(2, "mm"), type = "closed"),
    end_cap = circle(2, "mm"),
    strength = 0.5 # scales::rescale(g_actors$weight, 1, 4) # TODO fix this 
  ) +
  
  geom_node_point(
    aes(
      size = V(g_actors)$str_in,
      color = command_team
      ),
    alpha = 0.8
  ) +

  geom_node_label(
    aes(
      label = name
      ),
    repel = TRUE,
    size = 3,
    
  ) +

  # TODO fix scales
  scale_edge_color_manual(values = command_colors) + 
  scale_color_manual(values = command_colors) + 
  scale_size_continuous() +

  # TODO fix legends
  # legend
  guides(
    size = guide_legend(
      # nrow = 2, byrow = TRUE, 
      title = "Strength in"
    ),
    fill = guide_legend(
      title = "Command Team"
      # nrow = 1, byrow = TRUE, title = "Command Team",
      # override.aes = list(shape = shape_vals)
    ),
    color = guide_legend(
      title = " Command Team"
      # ncol = 2, byrow = TRUE,
    ),
    edge_color = guide_legend(
      title = " Command Team"
      # ncol = 2, byrow = TRUE,
    )
  ) +

  labs(
    title    = "Terrorist Cell",
    subtitle = "A network plot of a terrorist cell.",
  )
p
```

## 4.) `visNetwork`

```{r}
library(visNetwork)

```

As opposed to the ggplot framework where most of the plot information is taken care of within the plot code, visNetwork is generally done by taking care of most of the information outside of the plotting code. This is done by creating data frames for the nodes and edges separately and then adding attributes with special names to them. For example, a node or edge attribute of "title" will be the information for the tooltip which pops up on hover-over by the mouse. The "group" is specifically used for the legend.

```{r}
# Build nodes
nodes <- igraph::as_data_frame(g, what = "vertices") %>%
  tibble::rownames_to_column("id") %>%
  mutate(
    id    = as.character(id),
    label = media,
    value = audience.size,
    group = type.label
  )

# Shape mapping
node_types <- sort(unique(nodes$group))
shape_vals <- c("dot", "square", "triangle")[seq_along(node_types)]
shape_map  <- setNames(shape_vals, node_types)
nodes$shape <- shape_map[nodes$group]

# Node colors (background)
node_colors <- setNames(scales::hue_pal()(length(node_types)), node_types)
nodes$color.background <- node_colors[nodes$group]
nodes$color.border <- "#444444"

nodes$title <- paste("Media Name: ", nodes$media, "<br>Media Type: ", nodes$type.label, "<br>Audience Size: ", nodes$audience.size)

# Build edges
edges <- igraph::as_data_frame(g, what = "edges") %>%
  mutate(
    from   = as.character(from),
    to     = as.character(to),
    width  = scales::rescale(weight, to = c(0.5, 5)),
    arrows = "to",
    smooth = TRUE
  )

# Edge colors
edge_types <- sort(unique(edges$type))
edge_colors <- setNames(scales::hue_pal()(length(edge_types)), edge_types)
edges$color <- edge_colors[edges$type]

# Dash mapping (linetype analogue)
edges$dashes <- edges$type == edge_types[2]   # example: second type dashed

edges$title <- paste("Reference Type: ", edges$type)

```

There are disadvantages to storing data this way (for example, what if the title attribute was natural in the dataset to represent the a political title?) But, once the data is in, it becomes very easy to produce a really good plot automatically.

```{r}
vis <- visNetwork(nodes, edges,
    main    = "Media Networks Link to and Reference Each Other",
    submain = "A network plot of networks.",
    footer  = "Data from: https://kateto.net/network-visualization.",
    width = "100%"
    )%>%
  visNodes(scaling = list(min = 5, max = 30)) %>%
  visEdges(smooth = list(enabled = TRUE, type = "dynamic")) %>%
  visOptions(
    highlightNearest = TRUE,
    nodesIdSelection = TRUE,
    collapse = TRUE
  ) %>%
  visInteraction(
    dragNodes = TRUE,
    dragView  = TRUE,
    zoomView  = TRUE
  ) %>%
  visPhysics(stabilization = TRUE)

vis

```



## 5.)  Your Turn 

Reproduce the same graph as above (or close to it, showing the same information) using visNetwork. Make the tooltip display the individual's name and their role. Don't worry about a legend on this plot.

```{r}
# Build nodes

nodes <- igraph::as_data_frame(g_actors, what = "vertices") %>%
  tibble::rownames_to_column("id") %>%
  mutate(
    id    = as.character(id),
    title = paste0(
      "<div style='align:center; margin: 5px;'><span style='font-weight:800; font-size:16pt'>",
      name,
      "</span><br /><span style='font-size:8pt'>",
      Role,
      "</span></div>"
      )
  )

# Build edges
edges <- igraph::as_data_frame(g_actors, what = "edges") |>
  mutate(
    from   = as.character(from),
    to     = as.character(to),
    width  = scales::rescale(weight, to = c(0.5, 1.5)),
    arrows = "to",
    smooth = TRUE
  )

# Shape mapping
node_types <- sort(unique(nodes$command_team)) # 2 unique options
shape_vals <- c("dot", "square")[seq_along(node_types)]
shape_map <- setNames(shape_vals, node_types)

# nodes$shape <- shape_map[nodes$command_team] # TODO figure this out


# Node colors (background)
node_colors <- setNames(scales::hue_pal()(length(node_types)), node_types)
# nodes$color.background <- node_colors[nodes$command_team] # TODO problem right here, subsetting by logical
nodes$color.border <- "#444444"

# Edge colors
edge_types <- sort(unique(edges$command_team))
edge_colors <- setNames(scales::hue_pal()(length(edge_types)), edge_types)
# edges$color <- edge_colors[edges$command_team] # TODO similar issue to above... wtf do I need to add which() to pull out indices, then pass it in?

# Dash mapping (linetype analogue)
edges$dashes <- !edges$command_team # already a logical

```


```{r}
set.seed(1)
vis <- visNetwork(nodes, edges,
                  main = "Terrorist Cell",
                  submain = "",
                  footer = "",
                  width = "100%"
  ) |>
  visNodes(scaling = list(min = 5, max = 30),
           font = list(size = 24)
           ) |>
  visEdges(smooth = list(enabled = TRUE, type = "dynamic")) |>
  visOptions(
    highlightNearest = TRUE,
    nodesIdSelection = TRUE,
    collapse = TRUE
  ) |>
  visInteraction(
    dragNodes = TRUE,
    dragView  = TRUE,
    zoomView  = TRUE
  ) |>
  visPhysics(stabilization = TRUE)

vis

```

#### Something to think about.

How centralized is this network? Are there any key players that could be removed to make the network fail in their goals? Centralized networks are highly resilient but also insecure, which can play into strategies when analyzing these networks. We'll talk about a number of measures of network structure next week.

```{r degree_centralization_time}
centr_degree(g, mode = "all")$centralization

```


```{r centerness_centralization_time}
centr_clo(g, mode = "all")$centralization

```


```{r betweenness_centralization_time}
centr_betw(g)$centralization

```


```{r eigen_centralization_time}
centr_eigen(g)$centralization

```

**Answer:**
Relatively low degree centralization means overall lower density and number of features like triads and other graphlets. Closeness centralization being moderate would indicate that a fair about of nodes are clumped together and not being far from one another in the network. Betweenness centralization being moderate would indicate that our network is relatively close knit, as a fair proportion of the nodes are on shortest paths compared to one another. Interestingly, the eigen centralization for the network is decently high and would indicate that our network has a number of prestigious nodes, near other prestigious nodes.

Samudra appears to be the biggest player in this network. It would fragment significantly if the network lost him.

## 6.) Results

What you need to do next is to "knit" your document into an html using the "Knit" button near the top of the screen. This will give you an HTML that you can turn in.
