---
title: "STAT332 Statistical Analysis of Networks Sec 01 - Lab 6 - Prof. Joe Reid"
author: "Andrew Sparkes"
date: "`r Sys.Date()`"
due: "27-02-2026"
output:
  html_document:
    self_contained: true
    df_print: paged
    code_folding: hide
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(gridExtra)
library(ggraph)

library(sna)
library(data.table)
library(gt)
library(igraph) 

source("../graph.R")
```

```{r user_configurable_settings}
rds_path <- "dolphins.rds"
```

# Your turn

---

The dolphins dataset is stored as a .rds object (igraph in r) same as
this dataset was. This is a network from observations of dolphins and is
a pretty fun one to consider for this task. It is unweighted,
undirected, and pretty straight forward for this exercise.

Please walk back through this lab and see what it shows you about node
and edge level measures for this community (or communities) of dolphins.
Identify important dolphins using the measures we have defined. Is it
always the same dolphin? What does this say about their roles within the
pod? How many communities are there and how do members play brokerage
roles in these communities?

## Setup

```{r load_the_dataset}
g <- readRDS(rds_path)

summary(g)
```

So 62 nodes with 159 edges, undirected, and not weighted. Again, lets use a table summary for finding out some other basic properties:

```{r}
# ID = c("Order", "Size", "Connected", "Directed", "Acyclic", "Weighted", "Simple", "Has Loops", "Is Multigraph", "Bipartite", "Tree", "Forest")
# 
# Value = c(gorder(g), ecount(g), is_connected(g), is_directed(g), is_acyclic(g), is_weighted(g), is_simple(g), any_loop(g), any_multiple(g), is_bipartite(g), is_tree(g), is_forest(g))
# 
# table <- data.table(ID, Value)
# table %>% gt() %>% fmt_tf(columns = Value, tf_style = "yes-no")

print_big_fancy_summary_table(g) 
```
Ugly table, but the graph is connected and simple, as mentioned above! Excellent.

I think that's enough for the setup, so let's move on to our metrics:

# Micro-scale Measures

---

## Articulation Points and Bridges

### Weak Articulation

Weak Articulation points are easy to calculate. These are the ones that
are built into igraph.

```{r}
weak_ap <- articulation_points(g)
weak_ap

```

### Strong Articulation

```{r}
strong_ap <- strong_articulation_points(g) # function definition moved to graph.R
strong_ap

```

### Bridges

```{r}
bridges_g <- bridges(g)

E(g)[bridges_g]
```

```{r}
strong_bridges <- strong_bridges(g) # function definition moved to graph.R
strong_bridges

```

## Degree Measures

### In-Degree

```{r}
deg_in  <- degree(g, mode = "in")
deg_in

summary(deg_in)

plot_dist(deg_in)

V(g)$deg_in <- deg_in
```

### Out-Degree

```{r}
deg_out <- degree(g, mode = "out")
deg_out

summary(deg_out)

plot_dist(deg_out)

V(g)$deg_out <- deg_out
```

## Centrality Measures

### Normalized In-Degree Centrality

```{r}
deg_in_cent <- degree(g, v = V(g), mode = "in", loops = TRUE, normalized = TRUE)
deg_in_cent

summary(deg_in_cent)

plot_dist(deg_in_cent)

V(g)$deg_in_cent <- deg_in_cent
```

```{r}
deg_out_cent <- degree(g, v = V(g), mode = "out", loops = TRUE, normalized = TRUE)
deg_out_cent

summary(deg_out_cent)

plot_dist(deg_out_cent)

V(g)$deg_out_cent <- deg_out_cent
```

### Normalized Out-Degree Centrality

### K-Core Index

```{r}
in_core <- coreness(g, mode = "in")
in_core

summary(in_core)

plot_dist(in_core)

V(g)$in_core <- in_core
```

```{r}
out_core <- coreness(g, mode = "out")
out_core
summary(out_core)

plot_dist(out_core)

V(g)$out_core <- out_core
```

### Normalized Betweenness Centrality

```{r}
btw_cent <- betweenness(g, directed = TRUE, normalized = TRUE)
btw_cent

summary(btw_cent)

plot_dist(btw_cent)

V(g)$btw_cent <- btw_cent
```

### Normalized In-Closeness Centrality

```{r}
clo_in  <- closeness(g, mode = "in",  normalized = TRUE)
clo_in

summary(clo_in)

plot_dist(clo_in)

V(g)$clo_in <- clo_in
```

### Normalized Out-Closeness Centrality

```{r}
clo_out  <- closeness(g, mode = "out",  normalized = TRUE)
clo_out

summary(clo_out)

plot_dist(clo_out)

V(g)$clo_out <- clo_out
```

### Eigenvector - Not for directed graphs
N/A

### Page Rank

```{r}
pr <- page_rank(g, directed = TRUE)$vector
pr

summary(pr)

plot_dist(pr)

V(g)$pagerank <- pr
```

### Bonacich Centrality

```{r}
bon <- power_centrality(g, exponent = 0.1, rescale = TRUE)
bon

summary(bon)

plot_dist(bon)

V(g)$bonacich <- bon
```

### Edge Betweenness Centrality

```{r}
ebtw <- edge_betweenness(g, directed = TRUE)
# not printing this one out.  There are 400+ edges here.
# Also, there is no "normalized" version of edge betweenness in igraph
summary(ebtw)

E(g)$edge_betweenness <- ebtw
```

### Nearest Neighbor Edge Centrality

```{r}
deg_in  <- degree(g, mode = "in")
deg_out <- degree(g, mode = "out")
deg_tot <- deg_in + deg_out


E(g)$nne <- nne_edge_centrality(g, deg_tot) # function definition moved to graph.R

summary(E(g)$nne)
```

### Nearest Neighbor Edge In Centrality

```{r}
E(g)$nne_in <- sapply(E(g), function(e) {
  v <- ends(g, e)
  (deg_in[v[1]] - 1) + (deg_in[v[2]] - 1)
})

summary(E(g)$nne_in)
```

### Nearest Neighbor Edge Out Centrality

```{r}
E(g)$nne_out <- sapply(E(g), function(e) {
  v <- ends(g, e)
  (deg_out[v[1]] - 1) + (deg_out[v[2]] - 1)
})

summary(E(g)$nne_out)
```

### Local Clustering Coefficient

```{r}
cc_local <- transitivity(as.undirected(g, mode = "collapse"), type = "local", isolates = "zero")
cc_local

summary(cc_local)

plot_dist(cc_local)

V(g)$clust_local <- cc_local
```

### Local Efficiency

```{r}
le_in <- local_efficiency(g, directed = TRUE, mode = "in")
le_in

summary(le_in)

plot_dist(le_in)

V(g)$local_eff_in <- le_in
```

```{r}
le_out <- local_efficiency(g, directed = TRUE, mode = "out")
le_out

summary(le_out)

plot(le_out)

V(g)$local_eff_out <- le_out
```

### Burt's Constraint

```{r}
bc <- constraint(g)
bc

summary(bc)

plot(bc)

V(g)$constraint <- bc
```

## Roles ---

### Communities

This is a directed graph, so I'm going to use infomap. Technically, you
can use other methods that support directed graphs as well, but
brokerage depends on community membership.

```{r}
comm <- cluster_infomap(g)
V(g)$community <- membership(comm)
sizes(comm)

```

# Next Section

---

## SNA

We finally have to leave the world of igraph. The library SNA (social
network analysis) has the functions we are looking for

```{r}
A <- as.matrix(as_adjacency_matrix(g, sparse = FALSE))

grp <- V(g)$community
grp2 <- V(g)$shape
```

## Brokerage

```{r}
brok <- brokerage(A, cl = grp)
```

```{r}
summary(brok)
```

LOL. Right. So, now we have something to figure out. WHAT does this all
mean? w_I is a coordinator w_0 is an itinerant broker b_I0 is a
gatekeeper b_0I is a representative b_0 is a liasion t is combined.

The z-scores are calculated as:

$$ z = \frac{\text{observed brokerage} - \text{expected brokerage}}{\text{SD under random null}}$$

So a positive z indicates more brokerage than expected and a negative z
indicates less brokerage than expected.\
But these z scores need to be taken with a grain of salt and compared
against the actual graph visualization. Look for very high scores
relative to other scores.\
For example: V4 has a W_I z score of more than 13. This indicates that
there is pretty strong evidence that V4 is a coordinator.\
LIP seems to have a very high b_I0 score indicating that it may serve as
a gate keeper.\
VIP seems to act as both a representative and a gate keeper in cluster
2.

Now, remember that there were natural groups already identified in this
network? Let's try to use those instead (they are stored in grp2)

```{r, eval=FALSE}
# TODO ASK JOE this cell kept breaking and I was unable to determine why
brok2 <- brokerage(A, cl = grp2)

summary(brok2)
```

## Prestige

Obviously in-degree is a measure of prestige. The other methods are page
rank, authority scores, and hub scores.

```{r}

auth <- hits_scores(g, scale = TRUE)
auth$hub
auth$authority
```

```{r}
plot_dist(auth$authority)
```

```{r}
plot_dist(auth$hub)
```

```{r}
V(g)$authority <- auth$authority
V(g)$hub <- auth$hub
```

# TODO COPY LEIDEN GRAPH TO ABOVE AND CUT OFF RIGHT HERE AND DELETE REMAINDER BEFORE TURN IN

# One last thing.

---

We've been saving all of these measures as node and edge properties...

```{r}
summary(g)
```

TODO Perhaps run it through leiden and note community measure stuff?

```{r}
leiden_g <- cluster_leiden(g,
                           objective_function = "modularity",
                           weights = NA,
                           resolution = 1,
                           n_iterations = 6,
                           vertex_weights = NULL)
plot(leiden_g, g)
```


```{r}
p <- ggraph::ggraph(g, layout = "fr") +

  geom_node_text(
    aes(
      label = id
      ),
    repel = TRUE,
    size = 3
  ) +

# edge settings
  geom_edge_fan(
    aes(
      color = "#ff00ff"
      ),
    alpha = 0.8,
    lineend = "round",
    arrow = arrow(length = unit(0, "mm"), type = "closed"),
    end_cap = circle(0, "mm")
  ) +
  
# vertex settings
  geom_node_point(
    aes(
      size = 10 + btw_cent #,
      # color = community
      ),
    alpha = 0.6
  ) +

# legends
  
  # scale_size_continuous(name = "Age", limits = range(V(g)$age), breaks = range(V(g)$age)) +

# guides
  guides(
    country = guide_legend(),
    hand = guide_legend(),
    size = guide_legend(nrow = 2, byrow = TRUE),
    fill = guide_legend(
      nrow = 1, byrow = TRUE, title = "Vertex Color",
      # override.aes = list(fill = color)
    )
  ) +

# titles and text
# - title top center: "Tennis Losses from 2007", color black, bolded
# - subtitle top center (below title): "Data represents players from USA, GBR, and AUS", color black, normal font size
# - caption bottom center: "Data from networkdata R package", wta dataset, color gray, small
  labs(
    title    = "Tennis Losses from 2007",
    subtitle = "Data represents players from USA, GBR, and AUS"
  ) +
  theme_graph()
p

```


Identify important dolphins using the measures we have defined. Is it
always the same dolphin? What does this say about their roles within the
pod? How many communities are there and how do members play brokerage
roles in these communities?

**Answer**: 

